---
title: "Prepare Climate data"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set( 
                      warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse) |> suppressPackageStartupMessages()
library(sf)
library(terra)
library(purrr)
```

### Prepare daily data 

```{r}
# Create a custom function to extract the data
extract_climate_data <- function(shapefile, raster_folder_path) {

  shapefile_to_extract <- terra::vect(shapefile)

  # Read the raster files
  files <- list.files(path = raster_folder_path, pattern = ".tif$", full.names = TRUE)

  # Read a raster to compare crs
  r <- terra::rast(files[1])

  # Check projection and transform (project)
  if (crs(r) != crs(shapefile_to_extract)) {
    shapefile_to_extract <- terra::project(shapefile_to_extract, r)
  }

  # Extract all data
  myF <- function(x) {
    terra::extract(terra::rast(x), shapefile_to_extract, xy = TRUE)
  }

  extracted_data <- files |>
    map(myF) |>
    reduce(inner_join)

  return(extracted_data)

}
```

```{r, eval=FALSE}
geo_pinos  <- "data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp"

tmed_path <- "/Users/ajpelu/Desktop/DATA/CLIMA/REDIAM/02_DIARIA/TEMP_MEDIA_DIARIA/InfGeografica/InfRaster/COG/"
tmin_path <- "/Users/ajpelu/Desktop/DATA/CLIMA/REDIAM/02_DIARIA/TEMP_MINIMA_DIARIA/InfGeografica/InfRaster/COG/"
tmax_path <- "/Users/ajpelu/Desktop/DATA/CLIMA/REDIAM/02_DIARIA/TEMP_MAXIMA_DIARIA/InfGeografica/InfRaster/COG/"
prec_path <- "/Users/ajpelu/Desktop/DATA/CLIMA/REDIAM/02_DIARIA/PRECIP_DIARIA/InfGeografica/InfRaster/COG/"

startTime <- Sys.time() 

tmed <- extract_climate_data(geo_pinos, raster_folder_path = tmed_path)
startTime <- Sys.time() 
tmin <- extract_climate_data(geo_pinos, raster_folder_path = tmin_path)
tmax <- extract_climate_data(geo_pinos, raster_folder_path = tmax_path)
prec <- extract_climate_data(geo_pinos, raster_folder_path = prec_path)

endTime <- Sys.time()  
print(endTime - startTime)

write_csv(prec, "data/raw/climate/daily_prec.csv")
write_csv(tmed, "data/raw/climate/daily_tmed.csv")
write_csv(tmax, "data/raw/climate/daily_tmax.csv")
write_csv(tmin, "data/raw/climate/daily_tmin.csv")

saveRDS(prec, file = "data/raw/climate/daily_prec.rds")
saveRDS(tmed, file = "data/raw/climate/daily_tmed.rds")
saveRDS(tmax, file = "data/raw/climate/daily_tmax.rds")
saveRDS(tmin, file = "data/raw/climate/daily_tmin.rds")
```


```{r, eval=FALSE}
prepara_df <- function(x){
  out <- x |>
    pivot_longer(-c("ID","x", "y")) |>
    separate(name, into=c("var", "year", "month", "day", "cog")) |>
    dplyr::select(-cog)
  return(out)
}

mylist <- list(tmax, tmed, tmin, prec)

df <- mylist |>
  map(prepara_df) |>
  reduce(rbind) |> 
  mutate(var = case_when(
    var == "td1" ~ "tmed", 
    var == "td2" ~ "tmin",
    var == "td4" ~ "tmax", 
    var == "p" ~ "prec" 
  )) |> 
  mutate(
    year = as.numeric(year), 
    month = as.numeric(month), 
    day = as.numeric(day)
  )

geodf <- terra::vect(geo_pinos) |> as.data.frame() |>
  rownames_to_column(var = "ID") |>
  mutate(ID = as.numeric(ID)) |> 
  dplyr::select(ID, site_code, sp_code, elev_code)

climate <- df |> inner_join(geodf)

saveRDS(climate, "data/dendroadaptamed_climate_daily.rds")
```

### Compute the yearly and season data 

First, we compute the day of the year as a numeric value. Then, we apply a function called `date2season`, which categorizes each day into one of the four seasons according to the following distribution: 

- winter for days 1 to 80 and days 356 to 366, 
- spring for days 81 to 172, 
- summer for days 173 to 264, 
- autumn for days 265 to 355.

```{r}
climate_daily <- readRDS("data/dendroadaptamed_climate_daily.rds") |> 
  dplyr::select(-x,-y,-ID) |> 
  mutate(date = lubridate::make_date(year=year, month=month, day = day),
         yday = lubridate::yday(date)) |> 
  unite("sp_elev", c(sp_code, elev_code), remove = FALSE)

```


```{r}
date2season <- function(x) { 
  
  day_of_year <- lubridate::yday(x) 
  winter <- which(day_of_year %in% c(1:80,356:366))
  spring <- which(day_of_year %in% 81:172)
  summer <- which(day_of_year %in% 173:264)
  autumn <- which(day_of_year %in% 265:355)
  
  seasons <- rep(NA, length(x))
  seasons[winter] <- "winter"
  seasons[spring] <- "spring"
  seasons[summer] <- "summer"
  seasons[autumn] <- "autumn"
  
  return(seasons)
  
}

```

- Compute the season and the hydrological year (added as variables)

```{r}
climate <- climate_daily |> 
  mutate(season = date2season(date)) |> 
  mutate(month_names = strftime(date, '%b')) |> 
  mutate(
    hydro_month = case_when(
      month >= 10 ~ (month - 9), 
      TRUE ~ month + 3), 
    hydro_year = case_when(
      month >= 10 ~ (year + 1), 
      TRUE ~ year), 
    name_m = case_when(
      month >= 10 ~ paste0(month_names, "_pre"), 
      TRUE ~ month_names
    ))

```


- Compute the average values by year 

```{r}
avg_yearly <- bind_rows(
  # temp 
  climate |>
  filter(var != "prec") |> 
  group_by(var, sp_elev, year) |> 
  summarise(
    avg = mean(value, na.rm=TRUE),
    sd = sd(value, na.rm=TRUE),
    se = sd/sqrt(length(value))
  ),
  # prec
  climate |>
  filter(var == "prec") |> 
  group_by(var, sp_elev, year) |> 
  summarise(
    avg = sum(value, na.rm=TRUE),
    sd = NA,
    se = NA
  )
) |> separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE) |> 
  mutate(sp_code = as.factor(sp_code), 
         elev_code = as.factor(elev_code), 
         season = "yearly")

```


- Compute the average values by hydrological year ($Oct_{t-1} \sim Sep_{t+1}$). 


```{r}
avg_hidroyearly <- bind_rows(
  # temp 
  climate |>
  filter(var != "prec") |> 
  group_by(var, sp_elev, hydro_year) |> 
  summarise(
    avg = mean(value, na.rm=TRUE),
    sd = sd(value, na.rm=TRUE),
    se = sd/sqrt(length(value))
  ),
  # prec
  climate |>
  filter(var == "prec") |> 
  group_by(var, sp_elev, hydro_year) |> 
  summarise(
    avg = sum(value, na.rm=TRUE),
    sd = NA,
    se = NA
  )
) |> separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE) |> 
  mutate(sp_code = as.factor(sp_code), 
         elev_code = as.factor(elev_code), 
         season = "hydro year") |> 
  rename(year = hydro_year)
```

- Compute the average values by season 

```{r}
avg_season <- bind_rows(
  # temp 
 climate |>
  filter(var != "prec") |> 
  group_by(var, sp_elev, year, season) |> 
  summarise(
    avg = mean(value, na.rm=TRUE),
    sd = sd(value, na.rm=TRUE),
    se = sd/sqrt(length(value))
  ),
  # prec
  climate |>
  filter(var == "prec") |> 
  group_by(var, sp_elev, year, season) |> 
  summarise(
    avg = sum(value, na.rm=TRUE),
    sd = NA,
    se = NA
  )
) |> separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE) |> 
  mutate(sp_code = as.factor(sp_code), 
         elev_code = as.factor(elev_code))
```

- Join all the data, and store them as `data/dendroadaptamed_climate_season.rds`


```{r}
climate_avg <- 
  bind_rows(avg_yearly, 
            avg_hidroyearly, 
            avg_season)


saveRDS(climate_avg, "data/dendroadaptamed_climate_season.rds")
```


```{r, eval=FALSE}
climate_avg |> group_by(sp_elev, var, season) |> 
  summarise(min = min(year), max = max(year), n = length(year)) |> 
  unite(col = "range", min:max, sep = "-") |> 
  mutate(range = paste0("(", range, ")")) |> 
  unite(col = "s", n:range, sep = " ") |> 
  pivot_wider(values_from = s, names_from = season)
```








