---
title: "Climate characterization"
output: workflowr::wflow_html
bibliography: references.bib
cls: ecology-letters.csl
editor_options:
  chunk_output_type: console
---

# Introduction

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10, fig.height = 7
)
```


## Characterization of climate profile for each site

```{r}
library(tidyverse)
library(purrr)
library(multcompView)
library(DT)
library(parameters)
library(modelbased)
library(kableExtra)
library(emmeans)

source("scripts/aux.R")
```

```{r read-data}
climate <- readRDS("data/dendroadaptamed_climate_daily.rds") |> 
  dplyr::select(-x,-y,-ID) |> 
  mutate(date = lubridate::make_date(year=year, month=month, day = day),
         yday = lubridate::yday(date)) |> 
  unite("sp_elev", c(sp_code, elev_code), remove = FALSE)

```

- Data exploration by site. Creation of climatic diagrams
- To do this, we computed the average value for each day of the year throughout the time series (we refer to the dataset as `avg_daily`).
- We generated two versions of the climatic diagrams: one including the `low2` plot (Figure \@ref(fig:climate-diagram-all)) and another excluding it (Figure \@ref(fig:climate-diagram-nolow2)).


```{r prepare-climate-daily}
avg_daily <- climate |> 
  group_by(var, sp_elev, yday) |> 
  summarise(
    mean = mean(value, na.rm=TRUE),
    sd = sd(value, na.rm=TRUE),
    se = sd/sqrt(length(value))
  ) |> 
  mutate(date = as.Date(yday, origin = "2015-01-01", tz = "UTC") - 1) |> # convert to date. Be carefully because in R, the date index starts at 0. 2015 as "simulated year"
  ungroup() |> 
  separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE, sep = "_")
```

```{r}
# For add text with elevation 
geodf <- terra::vect( "data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp") |> 
  as.data.frame() |>
  mutate(elev_code = fct_recode(elev_code, `low-Dec` = "low2")) |> 
  mutate(elev_code = fct_relevel(elev_code, "low-Dec","low", "medium", "high")) |>
  unite("sp_elev", c(sp_code, elev_code), remove = FALSE) |> 
  dplyr::select(sp_elev, elev) |> 
  mutate(date = as.Date("2015-02-01"), 
          mean = 100) |> 
  separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE, sep = "_")

```



```{r climate-diagram-all, fig.cap="Climate diagrams for the study sites."}
# To get the scale factor needed for the secondary axis: max(15-days prec) / max(tmed) 
max_tmed <- avg_daily |> filter(var == "tmed") |> summarise(max(mean))


max_15prec <- avg_daily |>
         filter(var == "prec") |>
         mutate(m = lubridate::month(date)) |>
         group_by(sp_elev, m) |>
         summarise(mean = sum(mean)) |>
         mutate(date = make_date(year = "2015", month = m, day = "15")) |> 
  ungroup() |> 
  summarise(max(mean))

scale_factor <- max_15prec/max_tmed 
scale_factor <- 4.41



theme_climate_diagram <- function() {
  theme(
    strip.text.y = element_text(face = "italic"),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    axis.text.x = element_text(size = 6), 
    strip.background = element_rect(fill = "white")
  )
}


climate_diagram <- ggplot(
  data = (avg_daily |> filter(var == "tmed")), 
  aes(x = date, y = mean*scale_factor)
  ) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") + 
  geom_bar(
    data =
      (avg_daily |>
         filter(var == "prec") |>
         mutate(m = lubridate::month(date)) |>
         group_by(sp_elev, m) |>
         summarise(mean = sum(mean)) |>
         mutate(date = make_date(year = "2015", month = m, day = "15")) |> 
         separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE, sep = "_")
      ),
      aes(y = mean, fill = sp_code), stat = "identity"
  ) +
  geom_ribbon(aes(ymin = scale_factor*(mean - sd), 
                  ymax = scale_factor*(mean + sd), 
                  fill = sp_code), alpha=.5) + 
  geom_line() +
  ggh4x::facet_grid2(sp_code~factor(elev_code, levels=c('low', 'low-Dec', 'medium', 'high')), 
                     render_empty = FALSE) +
  scale_y_continuous("Precipitation (mm)",
                     sec.axis = sec_axis(~ ./scale_factor, name = "Mean Temperature (°C)")
  ) + 
  scale_fill_manual(values = colours_sp, name = "Species", guide = "none") +
  xlab("") + 
  theme_bw() + 
  theme_climate_diagram() +
  geom_text(data = geodf, aes(label = paste(elev, "m"), y = Inf), vjust = 1.5)


ggsave(
  climate_diagram, 
  file = "output/climate_diagram.png",
  dpi = 300,
  width = 12, height = 9
)

climate_diagram
```


```{r climate-diagram-nolow2,  fig.cap="Climate diagrams for the study sites except the low2 plot (P. pinaster)."}
climate_diagram_nolow2 <- ggplot(
    data = (avg_daily |> filter(elev_code != "low-Dec") |> filter(var == "tmed")), 
      aes(x = date, y = mean*scale_factor)
    ) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") + 
  geom_bar(
    data =
      (avg_daily |>
         filter(var == "prec") |>
         filter(elev_code != "low-Dec") |> 
         mutate(m = lubridate::month(date)) |>
         group_by(sp_elev, m) |>
         summarise(mean = sum(mean)) |>
         mutate(date = make_date(year = "2015", month = m, day = "15")) |> 
         separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE, sep = "_")
      ),
      aes(y = mean, fill = sp_code), stat = "identity"
  ) +
  geom_ribbon(aes(ymin = scale_factor*(mean - sd), ymax = scale_factor*(mean + sd), fill = sp_code), alpha=.5) + 
  geom_line() +
  ggh4x::facet_grid2(sp_code~factor(elev_code, levels=c('low','medium', 'high')), render_empty = FALSE) +
  scale_y_continuous("Precipitation (mm)",
                     sec.axis = sec_axis(~ ./scale_factor, name = "Mean Temperature (°C)")
  ) + 
  theme_bw() + 
  theme_climate_diagram() + 
  xlab("") + 
  geom_text(data = (geodf |> filter(elev_code != "low-Dec")), 
            aes(label = paste(elev, "m"), y = Inf), vjust = 1.5) +
  scale_fill_manual(values = colours_sp, name = "Species", guide = "none") 


climate_diagram_nolow2

ggsave(
  climate_diagram_nolow2, 
  file = "output/climate_diagram_nolow2.png",
  dpi = 300,
  width = 12, height = 9
)
```

- Anohther variant could be to include the max and minimum data (Figure \@ref(fig:climate-diagram-minmax)).

```{r climate-diagram-minmax,  fig.cap="Climate diagrams for the study sites except"}
# If we want to incorporate the max and minimum data 
climate_diagram + 
  geom_line(data = (avg_daily |> filter(var == "tmax")), 
            aes(x = date, y = mean*scale_factor), colour="red") +
  geom_ribbon(data = (avg_daily |> filter(var == "tmax")),
              aes(ymin = scale_factor*(mean - sd), ymax = scale_factor*(mean + sd)), 
              fill = "gray", alpha=.5) +
  geom_line(data = (avg_daily |> filter(var == "tmin")), 
            aes(x = date, y = mean*scale_factor), colour="blue") +
  geom_ribbon(data = (avg_daily |> filter(var == "tmin")),
              aes(ymin = scale_factor*(mean - sd), ymax = scale_factor*(mean + sd)), 
              fill = "gray", alpha=.5) 
```


## Differences among sites 

We explore the differences among sites yearly and seasonally. For this, we load the season and yearly data (see [preparation of climate data](climate_prepare.html)) 

```{r}
avg_seasonally_raw <- readRDS("data/dendroadaptamed_climate_season.rds")
```

### Climate characteristic of the sites 

```{r}
aux <- avg_seasonally_raw |> 
  filter(season == "yearly") |> 
  dplyr::select(-sd, -se) |> 
  pivot_wider(names_from = var, values_from = avg)

# aparecen NAs porque no hay datos de temp para los años 1951 a 1971 

temp_prec_sp <- aux |> 
  ggplot(aes(y=tmed, x = prec, colour = sp_code, 
             shape=elev_code)) +
  geom_point(size = 1.7, alpha = .8) + 
  facet_wrap(~factor(sp_code, levels = c("halepensis", "pinaster", "nigra", "sylvestris")), nrow = 1) +
  scale_colour_manual(values = colours_sp, name = "Species") +
  scale_shape_manual(values = shape_elev, name = "Elevation") +
  xlab("Annual precipitation (mm)") + 
  ylab("Average annual mean temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), 
        strip.text.x = element_text(face = "italic"), 
        strip.background = element_rect(fill = "white")) +
  scale_x_continuous(limits = c(0,2000))

ggsave(
  temp_prec_sp, 
  file = "output/climate_temp_prec_sp.png",
  dpi = 300,
  width = 8, height = 4
)

temp_prec_sp
```

```{r}
temp_prec_sp_elev  <- aux |> 
  ggplot(aes(y=tmed, x = prec, colour = sp_code, 
             shape=elev_code)) +
  geom_point(size = 1.5, alpha = .8) + 
  facet_grid(factor(sp_code, levels = c("halepensis", "pinaster", "nigra", "sylvestris")) ~ factor(elev_code, levels=c('low-Dec', 'low','medium', 'high'))) +
  scale_colour_manual(values = colours_sp, name = "Species") +
  scale_shape_manual(values = shape_elev, name = "Elevation") +
  xlab("Annual precipitation (mm)") + 
  ylab("Average annual mean temperature (ºC)") +
  theme_bw() +
  theme(legend.position = "none", 
        panel.grid.minor = element_blank(), 
        strip.text.y = element_text(face = "italic"), 
        strip.background = element_rect(fill = "white")) +
  scale_x_continuous(limits = c(0,2000))

ggsave(
  temp_prec_sp_elev, 
  file = "output/climate_temp_prec_sp_elev.png",
  dpi = 300,
  width = 6, height = 6
)

temp_prec_sp_elev 
```


### Comparisons among sites 

- Compute statistical comparisons among sites for `tmed` and `prec`.
- We compared the differences among elev, among sps and also within sps. 
- we applied a parametric tests (anova)

```{r}
# Function to generate comparisons and perform letter-based grouping
generate_comparisons <- function(model, sp_code = NULL, elev_code = NULL, alpha = 0.001) {
  emm <- emmeans(model, pairwise ~ sp_code * elev_code, adjust = "bonferroni", alpha = alpha,
                 at = list(sp_code = sp_code, elev_code = elev_code))
  
  
  cld_df <- multcomp::cld(emm, Letters = letters, level = 1 - alpha)
  p_values <- summary(emm, infer = TRUE)$contrasts$'p.value'
  result_df <- cld_df |> 
    mutate(p_value = p_values[match(.group, cld_df$.group)])
  return(result_df)
}
```



```{r}
tmed_yearly <- avg_seasonally_raw |> filter(season == "yearly") |> filter(var == "tmed")
alpha <- 0.001
lm_tmed<- lm(avg ~ sp_code*elev_code, data = tmed_yearly) 

means_letters_tmed <- bind_rows(
  generate_comparisons(lm_tmed, sp_code = "halepensis", elev_code = c("low", "medium", "high")),
  generate_comparisons(lm_tmed, sp_code = "pinaster", elev_code = c("low", "low-Dec", "medium", "high")),
  generate_comparisons(lm_tmed, sp_code = "nigra", elev_code = c("low", "medium", "high")),
  generate_comparisons(lm_tmed, sp_code = "sylvestris", elev_code = c("low", "medium", "high"))
)
```

```{r}
means_letters_tmed  |> 
  mutate(pvalue = ifelse(p_value < 0.0001, "<0.0001", round(p_value, 3))) |> 
  dplyr::select(-p_value) |> 
  arrange(sp_code, elev_code) |> 
  kbl(digits = c(0,0,2,2,0,2,2,5,0), 
      caption = "Posthoc differences among species for Tmed") |> 
    kable_paper("hover", full_width = F)
```


```{r}
prec_yearly <- avg_seasonally_raw |> filter(season == "yearly") |> filter(var == "prec")
alpha <- 0.001
lm_prec <- lm(avg ~ sp_code*elev_code, data = prec_yearly) 

means_letters_prec <- bind_rows(
  generate_comparisons(lm_prec, sp_code = "halepensis", elev_code = c("low", "medium", "high")),
  generate_comparisons(lm_prec, sp_code = "pinaster", elev_code = c("low", "low-Dec", "medium", "high")),
  generate_comparisons(lm_prec, sp_code = "nigra", elev_code = c("low", "medium", "high")),
  generate_comparisons(lm_prec, sp_code = "sylvestris", elev_code = c("low", "medium", "high"))
)
```

```{r}
means_letters_prec  |> 
  mutate(pvalue = ifelse(p_value < 0.0001, "<0.0001", round(p_value, 3))) |> 
  dplyr::select(-p_value) |> 
  kbl(digits = c(0,0,2,2,0,2,2,0,3), 
      caption = "Posthoc differences among species for Annual Rainfall") |> 
    kable_paper("hover", full_width = F)
```

- By elevation


```{r}
by_elevation <- bind_rows(
(emmeans(lm_tmed, specs = pairwise~elev_code) |> 
      multcomp::cld(Letters=letters, level=1-alpha) |> 
  mutate(var = "Tmed") |>
    relocate(var, .before = elev_code)),
(emmeans(lm_prec, specs = pairwise~elev_code) |> 
      multcomp::cld(Letters=letters, level=1-alpha) |> 
  mutate(var = "Prec") |>
    relocate(var, .before = elev_code))
)

```

```{r}
by_elevation  |> 
  kbl(digits = c(0,0,2,2,0,2,2,5,0), 
      caption = "Posthoc differences among elevation (not considering species) for Annual Rainfall and Tmed") |> 
    kable_paper("hover", full_width = F)
```


- By species 

```{r}
lm_tmed2 <- lm(avg ~ sp_code * elev_code,
  data = (tmed_yearly |> filter(elev_code != "low-Dec"))
)

lm_prec2 <- lm(avg ~ sp_code * elev_code,
  data = (prec_yearly |> filter(elev_code != "low-Dec"))
)

by_sp <- bind_rows(
  (emmeans(lm_tmed2, specs = pairwise ~ sp_code) |>
    multcomp::cld(Letters = letters, level = 1 - alpha) |>
    mutate(var = "Tmed") |>
    relocate(var, .before = sp_code)),
  (emmeans(lm_prec2, specs = pairwise ~ sp_code) |>
    multcomp::cld(Letters = letters, level = 1 - alpha) |>
    mutate(var = "Prec") |>
    relocate(var, .before = sp_code))
)
```

```{r}
by_sp  |> 
  kbl(digits = c(0,0,2,2,0,2,2,5,0), 
      caption = "Posthoc differences among species (not considering elevation) for Annual Rainfall and Tmed") |> 
    kable_paper("hover", full_width = F)
```


```{r}
library(ggpubr)

tmed_yearly |> 
  filter(elev_code != "low-Dec") |> 
  ggplot(aes(x=factor(elev_code, levels=c('low', 'low-Dec', 'medium', 'high')), y = avg)) +
  geom_violin() +
  geom_jitter()

 

ggboxplot(tmed_yearly, x = "elev_code", y = "avg", add = "jitter")


p <- ggboxplot(df, x = "dose", y = "len",
                color = "dose", palette =c("#00AFBB", "#E7B800", "#FC4E07"),
                add = "jitter", shape = "dose")
 p
```









```{r, eval=FALSE, echo=FALSE}

# 
# 
# 
# VAS POR AQUI MELÓN 
# #### Temporal trends
# - Plot the temporal trends (todo: compute the temporal trends statistics?)
# ```{r}
# avg_seasonally_raw |> 
#   filter(season == "yearly") |> 
# 
# 
# avg_yearly |> 
#   filter(var != "prec") |> 
#   filter(elev_code != "low2") |>
#   ggplot(aes(x = year, y = avg, colour=var)) +
#   geom_line() +
#   geom_line() +
#   facet_grid(sp_code~factor(elev_code, levels=c('low','medium', 'high'))) +
#   theme_bw() +
#   theme(
#     panel.grid.minor.x = element_blank(),
#     panel.grid.major.x = element_blank(), 
#     strip.text = element_text(size = 14)
#   ) +
#   xlab("Year") +
#   ylab("Temperature (yearly average) (°C)") +
#   ylim(c(0,25)) + 
#   scale_colour_manual(values = colours_temp)
# ```
# 
# ```{r}
# avg_yearly |> 
#   filter(var == "prec") |> 
#   filter(elev_code != "low2") |>
#   ggplot(aes(x = year, y = avg, colour=var)) +
#   geom_line() +
#   geom_line() +
#   facet_grid(sp_code~factor(elev_code, levels=c('low','medium', 'high'))) +
#   theme_bw() +
#   theme(
#     panel.grid.minor.x = element_blank(),
#     panel.grid.major.x = element_blank(), 
#     strip.text = element_text(size = 14)
#   ) + 
#   geom_smooth(method = "lm", se=FALSE, linewidth=.5) + 
#   xlab("Year") +
#   ylab("Precipitation (yearly average) (mm)") 
# ```
# 
# ### Bewteen sites comparison
# 
# 
# 
# 
# 
# ### Seasonal 
# 
# ```{r}
# avg_season <- bind_rows(
#   # temp 
#   d |>
#     filter(var != "prec") |> 
#     group_by(var, sp_elev, year, season) |> 
#     summarise(
#       avg = mean(value, na.rm=TRUE),
#       sd = sd(value, na.rm=TRUE),
#       se = sd/sqrt(length(value))
#     ),
#   # prec
#   d |>
#     filter(var == "prec") |> 
#     group_by(var, sp_elev, year, season) |> 
#     summarise(
#       avg = sum(value, na.rm=TRUE),
#       sd = NA,
#       se = NA
#     )
# ) |> separate(sp_elev, into = c("sp_code", "elev_code"), remove = FALSE) |> 
#   mutate(sp_code = as.factor(sp_code), 
#          elev_code = as.factor(elev_code))
# ```
# 
# - Plot the temporal trends (todo: compute the temporal trends statistics?)
# ```{r}
# plot_temp_season <- function(x, season_var, ycaption) { 
#   x |> 
#     filter(var != "prec") |> 
#     filter(season == season_var) |> 
#     filter(elev_code != "low2") |>
#     ggplot(aes(x = year, y = avg, colour=var)) +
#     geom_line() +
#     geom_line() +
#     facet_grid(sp_code~factor(elev_code, levels=c('low','medium', 'high'))) +
#     theme_bw() +
#     theme(
#       panel.grid.minor.x = element_blank(),
#       panel.grid.major.x = element_blank(), 
#       strip.text = element_text(size = 14)
#     ) +
#     xlab("Year") +
#     ylab(ycaption) +
#     scale_colour_manual(values = colours_temp)
#   
# }
# 
# plot_temp_season(avg_season, season_var = "spring", 
#                  ycaption = "Spring temperature (°C)") + ggtitle("Spring")
# 
# plot_temp_season(avg_season, season_var = "summer", 
#                  ycaption = "Summer temperature (°C)") + ggtitle("Summer")
# 
# plot_temp_season(avg_season, season_var = "autumn", 
#                  ycaption = "Autumn temperature (°C)") + ggtitle("Autumn")
# 
# plot_temp_season(avg_season, season_var = "winter", 
#                  ycaption = "Winter temperature (°C)") + ggtitle("Winter") 
# 
# ```
# 
# ```{r}
# plot_prec_season <- function(x, season_var, ycaption, line_colour="#67a9cf") { 
#   x |> 
#     filter(var == "prec") |> 
#     filter(season == season_var) |> 
#     filter(elev_code != "low2") |>
#     ggplot(aes(x = year, y = avg, colour=var)) +
#     geom_line() +
#     facet_grid(sp_code~factor(elev_code, levels=c('low','medium', 'high'))) +
#     theme_bw() +
#     theme(
#       panel.grid.minor.x = element_blank(),
#       panel.grid.major.x = element_blank(), 
#       strip.text = element_text(size = 14)
#     ) +
#     xlab("Year") +
#     ylab(ycaption) +
#     scale_colour_manual(values = line_colour)
#   
# }
# 
# plot_prec_season(avg_season, season_var = "spring", 
#                  ycaption = "Spring rainfall (mm)") + ggtitle("Spring")
# 
# plot_prec_season(avg_season, season_var = "summer", 
#                  ycaption = "Summer rainfall (mm)") + ggtitle("Summer")
# 
# plot_prec_season(avg_season, season_var = "autumn", 
#                  ycaption = "Autumn rainfall (mm)") + ggtitle("Autumn")
# 
# plot_prec_season(avg_season, season_var = "winter", 
#                  ycaption = "Winter rainfall (mm)") + ggtitle("Winter") 
# ```
# 
# 
# #### Bewteen sites comparison
# - Temperatures (ToDo ANOVA-2way or similar)
# 
# ```{r}
# 
# plot_compare_seasonal_temp <- function(x, season_var, ycaption){
#   x |> 
#     filter(var != "prec") |> 
#     filter(season == season_var) |> 
#     ggplot(aes(x = factor(sp_code, 
#                           levels = c("halepensis", "pinaster", "nigra", "sylvestris")),
#                y = avg, 
#                fill = factor(elev_code, levels = c("low2", "low", "medium", "high")))) +
#     geom_boxplot() + 
#     scale_fill_manual(values = colours_elev, name = "") +
#     theme_bw() +
#     theme(
#       legend.position = "bottom",
#       panel.grid = element_blank()
#     ) +
#     xlab("") +
#     ylab(ycaption) +
#     facet_wrap(~var, ncol = 1, scales= "free_y")
# }
# 
# 
# plot_compare_seasonal_temp(avg_season, season_var = "spring", 
#                            ycaption = "Spring temperature (°C)") + ggtitle("Spring")
# 
# plot_compare_seasonal_temp(avg_season, season_var = "summer", 
#                            ycaption = "Summer temperature (°C)") + ggtitle("Summer")
# 
# plot_compare_seasonal_temp(avg_season, season_var = "autumn", 
#                            ycaption = "Autumn temperature (°C)") + ggtitle("Autumn")
# 
# plot_compare_seasonal_temp(avg_season, season_var = "winter", 
#                            ycaption = "Winter temperature (°C)") + ggtitle("Winter")
# 
# ```
# 
# - Precipitation (ToDo ANOVA-2way or similar)
# 
# ```{r}
# avg_season |> 
#   filter(var == "prec") |> 
#   ggplot(aes(x = factor(sp_code, 
#                         levels = c("halepensis", "pinaster", "nigra", "sylvestris")),
#              y = avg, 
#              fill = factor(elev_code, levels = c("low2", "low", "medium", "high")))) +
#   geom_boxplot() + 
#   scale_fill_manual(values = colours_elev, name = "") +
#   theme_bw() +
#   theme(
#     legend.position = "bottom",
#     panel.grid = element_blank()
#   ) +
#   xlab("") +
#   ylab("Rainfall") +
#   facet_wrap(~season, ncol = 1, scales= "free_y")
# 
# ```
# 
# 
# 
# #### Tables
# 
# ```{r}
# avg_all <- rbind(avg_season, 
#                  avg_yearly |> mutate(season = "yearly"))
# 
# 
# summary_table <- avg_all |> 
#   group_by(var, sp_code, elev_code, season) |> 
#   summarise(
#     mean= round(mean(avg, na.rm=TRUE),2),
#     sd = round(sd(avg, na.rm=TRUE),2),
#     se = round(sd/sqrt(length(avg)),2)
#   ) |> 
#   unite("v", c("mean", "se"), sep = " ± ") |> 
#   dplyr::select(-sd) |> 
#   pivot_wider(names_from = var, 
#               values_from = v)
# ```
# 
# ```{r}
# summary_table |> 
#   DT::datatable(options =
#                   list(
#                     autoWidth = TRUE,
#                     pageLength = 15), 
#                 filter = list(position = 'top', clear = FALSE)
#                 
#   )
# 
# ```
# 
# 
# 
# ```{r}
# head(avg_yearly)
# 
# prec <- avg_yearly |> filter(var == "prec")
# 
# aov_prec <- aov(avg ~ sp_code*elev_code, data = prec) 
# 
# plot(aov_prec, which = 2)
# 
# car::qqPlot(aov_prec$residuals)
# 
# library(rstatix)
# 
# 
# aov_prec <- aov(annual_value ~ sp_code*elev_code, data = prec) 
# aov_tmed <- 
# 
# |> tukey_hsd()
# 
# gg |> filter(str_detect(group1, "halepensis")) |> 
#   filter(str_detect(group2, "halepensis"))
# 
# plot(predictorEffect(aov_prec))
# 
# plot(allEffects(aov_prec))
# with(prec, 
# interaction.plot(elev_code, sp_code, annual_value))
# ```
# gg

```




