---
title: "Compile database"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10, fig.height = 7
)
```


```{r}
library(tidyverse)
library(here)
library(tidyverse)
library(kableExtra)
library(DT)
library(hydroTSM)
library(DiagrammeR)
source("scripts/aux.R")
```


# Introduction

# Levels 

```{r, eval=FALSE}
geodf <- terra::vect( "data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp") |> 
  as.data.frame() |>
  filter(site_code != "PIHAL_NACIMIENTO_L") |> 
  filter(site_code != "PIHAL_NACIMIENTO_M") |> 
  filter(site_code != "PISYL_PORTUGOS_H") |> 
  mutate(site_code = str_remove(site_code, "modify")) |> 
  mutate(elev_code = case_when(
    elev_code == "low2" ~ "low-Dec", 
    TRUE ~ elev_code
  )) |> 
  unite("sp_elev", c(sp_code, elev_code), remove = FALSE) |> 
  dplyr::select(sp_elev, elev, site_code) |> 
  separate(sp_elev, into = c("sp_code", "elev_code"), sep = "_", remove = FALSE) |> 
  mutate(
    Specie = paste0("P. ", sp_code)
  )

write_csv(geodf, "data/levels_site.csv")
```

```{r}
#| results: asis
#| echo: FALSE

geodf <- read_csv("data/levels_site.csv")
geodf |> DT::datatable(rownames = FALSE)
```


# Remote Sensing Data

## IV 

### IV Modis 

- Features of MOD13Q1: pixel size = 250; temporal range = 2000 to 2021; 23 images per year
- Prepare data:
  
    - Select data for *Pinus* field plot
    - Extract date of each image
    - Format NDVI and EVI values (multiply for scale factor 0.0001; see this [link](https://developers.google.com/earth-engine/datasets/catalog/MODIS_006_MOD13Q1)) 

First, we generated a dataset with the following fields: 

```{r}
# Read raw data 
raw <- read_csv(here::here("data/raw/remote_sensing/iv_mod13q1_c61.csv"))

d_raw <- raw |>
  #filter(str_detect(Specie, "Pinu")) |>
  mutate(date = as.Date(str_replace_all(substr(`system:index`, 1, 10), "_", "-"),
                        format = "%Y-%m-%d"),
         evi = EVI*0.0001,
         ndvi = NDVI*0.0001) |> 
  dplyr::select(date, lat, long, site_code, evi, ndvi)

d <- inner_join(d_raw, geodf)
```

```{r}
#| echo: true
names(d)
```

Then, we compute 
- Compute statistics (*mean*, *sd*, *se*, *cv*, ...) for each IV indexes grouping by year/season for each field-plot. 
- We need to group by season and by year. 

```{r}
iv_yearly <- d |>
  mutate(year=lubridate::year(date)) |> 
  dplyr::select(site_code, Specie, year, evi, ndvi, sp_elev) |> 
  pivot_longer(cols = c("ndvi", "evi"), 
               names_to = 'variable', values_to = 'value') |> 
  group_by(site_code, Specie, year, sp_elev, variable) |> 
  summarise(mean = mean(value, na.rm = TRUE), 
            sd = sd(value, na.rm = TRUE), 
            se = sd/sqrt(length(value)),
            cv = sd/mean*100,
            median = median(value, na.rm=TRUE),
            sum = sum(value, na.rm=TRUE),
            min = min(value), 
            max = max(value),
            n = length(value)) |> 
  mutate(season = as.factor("yearly")) |> 
  ungroup() |> 
  separate(sp_elev, into = c("sp_code", "elev_code"), sep = "_", remove = FALSE)
```


```{r}
iv_season <- d |> 
  mutate(season = as.factor(hydroTSM::time2season(date, out.fmt="seasons"))) |> 
  mutate(year=lubridate::year(date)) |> 
  dplyr::select(site_code, Specie, year, season, evi, ndvi, sp_elev) |> 
  pivot_longer(cols = c("ndvi", "evi"), 
               names_to = 'variable', values_to = 'value') |> 
  group_by(site_code, Specie, year, sp_elev, season, variable) |> 
  summarise(mean = mean(value, na.rm = TRUE), 
            sd = sd(value, na.rm = TRUE), 
            se = sd/sqrt(length(value)),
            cv = sd/mean*100,
            median = median(value, na.rm=TRUE),
            sum = sum(value, na.rm=TRUE),
            min = min(value), 
            max = max(value),
            n = length(value)) |> 
  ungroup() |> 
  separate(sp_elev, into = c("sp_code", "elev_code"), sep = "_", remove = FALSE)
```

```{r}
write_csv(bind_rows(iv_season, iv_yearly), "data/dendroadaptamed_iv_modis.csv")
```

### Landsat 

```{r}
landsat <- read_csv("data/raw/remote_sensing/iv_landsat_c02.csv") |> 
  rename(evi = EVI, ndvi = NDVI, image = `system:index`) |> 
  dplyr::select(image, evi, ndvi, date, site_code) |> 
  mutate(date = as.Date(as.character(date), format="%Y%m%d")) |> 
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date)) |> 
  filter(year != 2023) |> 
  inner_join(geodf, by = "site_code")

evilandsat <- landsat |> 
  filter(!is.na(evi)) |> 
  group_by(sp_elev, sp_code, elev_code, year) |> 
  summarise(mean = mean(evi, na.rm = TRUE), 
            sd = sd(evi, na.rm = TRUE), 
            se = sd/sqrt(length(evi)),
            cv = sd/mean*100,
            median = median(evi, na.rm=TRUE),
            sum = sum(evi, na.rm=TRUE),
            min = min(evi), 
            max = max(evi),
            n = length(evi))

ndvilandsat <- landsat |> 
  filter(!is.na(ndvi)) |> 
  group_by(sp_elev, sp_code, elev_code, year) |> 
  summarise(mean = mean(ndvi, na.rm = TRUE), 
            sd = sd(ndvi, na.rm = TRUE), 
            se = sd/sqrt(length(ndvi)),
            cv = sd/mean*100,
            median = median(ndvi, na.rm=TRUE),
            sum = sum(ndvi, na.rm=TRUE),
            min = min(ndvi), 
            max = max(ndvi),
            n = length(ndvi))

iv_landsat <- bind_rows(
  (ndvilandsat |> mutate(iv = "ndvi")), 
  (evilandsat |> mutate(iv = "evi")))

```


```{r}
write_csv(iv_landsat, "data/dendroadaptamed_iv_landsat.csv")
```





### NPP 
- NPP data original as $kg ~ C ~ m^2 ~ year^{-1}$ but with a scale factor of 0.0001, so to transform to $g ~ C ~ m^2 ~ year^{-1}$ apply 0.1

- Npp_QC: quality control (percentage)

```{r prepare-data}
raw_npp <- read_csv(here::here("data/raw/remote_sensing/npp_mod17a3_yearly.csv"))

npp <- raw_npp |> 
  mutate(date = as.Date(str_replace_all(substr(`system:index`, 1, 10), "_", "-"),
                        format = "%Y-%m-%d"),
         npp = Npp*0.1) |> 
  mutate(sp_code = str_replace_all(Specie, "Pinus_", "")) |> 
  mutate(year = lubridate::year(date)) |> 
  dplyr::select(year, sp_code, elev_code, npp, npp_qc = Npp_QC) 

write_csv(npp, "data/dendroadaptamed_npp_mod17a3_yearly.csv")
```




## Tree ring Data 
### ABI data

```{r}
codes_files <- read_csv("data/codes_files_dendro.csv")
files_abi <- list.files("data/abi", full.names = TRUE)


custom_read_abi <- function(x) { 
  fnm <- gsub(pattern = ".csv", "" , basename(x))
  
  sp_elev <- codes_files |> 
    filter(abi_name %in% fnm)
  
  out <- read_csv(x) |> 
    mutate(abi_name = fnm) |> 
    inner_join(sp_elev) |> 
    dplyr::select(year = year1, IBT_ag_m2, sp_code, elev_code)
                
  return(out)
  }


abi <- files_abi |> purrr::map(custom_read_abi) |> bind_rows() 

write_csv(abi, "data/dendroadaptamed_abi.csv")
```


### BAI data 

```{r}
files_bai <- list.files("data/bai", full.names = TRUE)

custom_read_bai <- function(x) { 
  fnm <- gsub(pattern = ".csv", "" , basename(x))
  
  sp_elev <- codes_files |> 
    filter(bai_name %in% fnm)
  
  out <- read_csv(x) |> 
    mutate(bai_name = fnm) |> 
    inner_join(sp_elev) |> 
    dplyr::select(year, BAI, BAIserror, sp_code, elev_code)
                
  return(out)
  }


bai <- files_bai |> purrr::map(custom_read_bai) |> bind_rows() 

write_csv(bai, "data/dendroadaptamed_bai.csv")
```


