---
title: "Source Data"
output: workflowr::wflow_html
bibliography: references.bib
cls: ecology-letters.csl
editor_options:
  chunk_output_type: console
---


```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```



```{r pkg}
library(here)
library(tidyverse)
library(kableExtra)
library(DT)

source("scripts/aux.R")
```



# Data Sources

## Climatic data

### Daily data 

- To characterize the climatic conditions at each of the sampling sites, we utilized high-resolution climatic maps (spatial resolution of 500 meters) sourced from the Andalusian Information Subsystem for Environmental Climatology, part of the Andalusian Environmental Information Network. These maps encompassed dialy averages of mean, minimum, and maximum temperatures, in addition to dialy cumulative rainfall data. These detailed maps were constructed by integrating data from over 200 meteorological stations. For each individual sampling site, we extracted temporal data series spanning from 1951 and 1971 for rainfall and temperature series respectively. 

### SPEI

- The Drought indices dataset for Spain was used through the https://monitordesequia.aemet.es/ [@VicenteSerranoetal2017HighResolution]

- For each field site, we downloaded the time series using the web interface. It's not the most efficient way, but we don't want to download the entire `.nc` files and processing all (it's a time-consuming task and also used a lot of space). If we want to process a lot of localities, donwload the `.nc` files from https://monitordesequia.csic.es/historico 

-

# Satellite data 

## Vegetation Indexes 

- We extracted temporal series of vegetation indexes as a surrogate of productivity. Several data sources were used: 

### MODIS 
We used the MODIS product MOD13Q1 (temporal resolution: 16 days; spatial resolution = 250 m). We used NDVI and EVI spectral index. For each field plot, we extracted all the composite images of the MODIS product (using the centroid of the field plot). Then, a temporal series was created for each field plot. 
  
### Landsat Data 

- We used the Collection 2 of the Landsats products
- We combined [Landsat 5 TM](https://developers.google.com/earth-engine/datasets/catalog/LANDSAT_LT05_C02_T1_L2), [Landsat 7 ETM+](https://developers.google.com/earth-engine/datasets/catalog/LANDSAT_LE07_C02_T1_L2), and [Landsat 8 OLI](https://developers.google.com/earth-engine/datasets/catalog/LANDSAT_LC08_C02_T1_L2)
- For all datasets we used the Level 2 Data, Tier 1 (Surface Reflectance)
- We applied different masks at pixelwise. See [this](https://hendrik-wulf.gitbook.io/geo-717-earth-engine/day-2-time-series-analysis). For this purpose we applied a [CFMask Algorithm](https://www.usgs.gov/landsat-missions/cfmask-algorithm) that allow us to remove pixels flagged as clouds, cloud shadows, snow/ice, or water, allowing only clear pixels to remain for further analysis. See [this](https://doi.org/10.1016/j.rse.2017.03.026). 

## Annual Net Primary Productivity 

- MODIS [MOD17A3HGF](https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MOD17A3HGF)
- Spatial resolution: 500 m pixel
- The MOD17A3HGF V6.1 product provides information about annual Gross and Net Primary Productivity (GPP and NPP) at 500m pixel resolution. Annual NPP is derived from the sum of all 8-day Net Photosynthesis (PSN) products (MOD17A2H) from the given year. The PSN value is the difference of the Gross Primary Productivity (GPP) and the Maintenance Respiration (MR) (GPP-MR).
- Detailed info [here](https://lpdaac.usgs.gov/documents/972/MOD17_User_Guide_V61.pdf)

Data were obtained from GEE (see code) (TODO: improve code)

- We extracted temporal series of yearly NPP from MODIS product MOD17A3HGF (temporal resolution: yearly; spatial resolution = 500 m). See [here](https://lpdaac.usgs.gov/products/mod17a3hgfv061/) for a description of the method to obtain the NPP. 

### Quality Control 
- We used the quality control band (`NPP_QC`). 
- The values of this band range from 0 to 100. They indicate a percentage computed as the number of days during the growing season with filled MODIS LAI data due to contaminated data by the total number of days of the growing season. See section 4.2 of the [user guide](https://lpdaac.usgs.gov/documents/972/MOD17_User_Guide_V61.pdf)

```{r qc-npp}
#| fig.width: 8 
#| fig.height: 8
#| fig.cap: "Histogram of the quality values for NPP. The Npp_QC_500 % indicates the number of days filled due to contaminated data. Higher values indicated poorly data."
#| echo: false


dnpp <- read_csv("data/dendroadaptamed_npp_mod17a3_yearly_c61.csv") |> 
  filter(year <= 2022) 

q <- quantile(dnpp$npp_qc, probs = c(0.5, 0.75, 0.9))

dnpp |> ggplot(aes(x=npp_qc)) + geom_histogram() +
  geom_vline(xintercept = q[1]) +
  geom_vline(xintercept = q[2]) +
  geom_vline(xintercept = q[3]) +
  geom_label(aes(x = q[1], y = 50, label = paste0("Percentil\n 50%\n QC=", q[1]))) +
  geom_label(aes(x = q[2], y = 50, label = paste0("Percentil\n 75%\n QC=", q[2]))) +
  geom_label(aes(x = q[3], y = 50, label = paste0("Percentil\n 90%\n QC=", q[3]))) +
  theme_minimal() +
  xlab("Npp_QC_500m (%)") + 
  ylab("# pixels")

```



## References
