---
title: "Compare Drought events "
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```


```{r, message=FALSE}
library(tidyverse)
library(patchwork)
library(likelihood)
library(kableExtra)
library(patchwork)
source("scripts/aux.R")
```



```{r}
spei_data <- read_csv("data/dendroadaptamed_spei_climate.csv") 
```


## Compare the severe drought events. 
A severe drought event starts when SPEI falls below the threshold of −1.28 (Páscoa et al 2017; Spinoni et al. 2018). A drought event is considered only when SPEI values fall below that threshold for at least two consecutive months. For each drought event, we computed: 

- the duration as the number of consecutive months with the SPEI lower than a certain threshold
- the severity as the sum of the absolute SPEI values during the drought event
- the intensity and the Lowest SPEI refer to the mean and lowest value of SPEI, respectively, during the drought event.

We computed the severe drought events (below -1.28) by site and per each index (i.e. SPEI-06 and SPEI-12) since 1960. 

```{r}
droughtIndicators <- function(df, vname, threshold) {

  require(data.table)

  # Add a new column indicating if the value is below the threshold and the following month is also below the threshold
  out <- df |>
    mutate(is_drought = ifelse(
      .data[[vname]] < threshold & lead(.data[[vname]], default = .data[[vname]][n()]) < threshold, 1, 0
    )) |>
    mutate(date = lubridate::make_date(year, month))
  
  # Compute the drought duration of the events
  out2 <- out |>
    group_by(index_events = data.table::rleid(is_drought)) |>
    mutate(drought_duration = sum(is_drought)) |>
    as.data.frame()

  # Filter events with drought duration > 1
  out3 <- out2 |>
    filter(drought_duration > 1) |>
    as.data.frame()

  # Compute several indicators (drought assessments)
  da <- out3 |>
    group_by(index_events) |>
    summarise(
      d_duration = unique(drought_duration),
      d_intensity = mean(.data[[vname]], na.rm = TRUE),
      d_severity = sum(abs(.data[[vname]]), na.rm = TRUE),
      lowest_spei = min(.data[[vname]]),
      month_peak = month[which.min(.data[[vname]])],
      minyear = min(year),
      maxyear = max(year),
      rangeDate = paste(
        lubridate::month(min(date), label = TRUE), "-",
        (lubridate::month(max(date), label = TRUE))
      )
    ) |>
    as.data.frame()

  return(list(data = out2, drought_events = out3, drought_assessment = da))
}

```

```{r}
s <- spei_data |> 
  dplyr::select(sp_elev, year, month, spei06, spei12) |> 
  pivot_longer(c(spei06, spei12), names_to = "spei") |> 
  filter(!is.na(value))


spei_events <- s |> 
  group_by(spei, sp_elev) |> 
  nest() |> 
  mutate(drought_results = map(data, ~droughtIndicators(., "value", -1.28)$drought_assessment)) |> 
  unnest(drought_results) |> 
  select(-data)
```


```{r}
d <- spei_events |> 
  separate(rangeDate, into = c("minmonth", "maxmonth"), remove = FALSE) |> 
  unite("start_date", minyear, minmonth, sep="-", remove = FALSE) |> 
  unite("end_date", maxyear, maxmonth,  sep="-", remove = FALSE) |> 
  mutate(start_date = as.Date(paste0(start_date, "-01"), format="%Y-%b-%d")) |> 
  mutate(end_date = as.Date(paste0(end_date, "-01"), format="%Y-%b-%d")) |> 
  rowwise() |> 
  mutate(middate = (start_date + ((end_date - start_date)/2))) 
```


#### Drought events SPEI-12 

```{r}
#| fig.width: 10
#| fig.height: 12
d12 <- d |> 
  filter(spei == "spei12") |> 
  ungroup() |> 
  dplyr::select(sp_elev, d_duration, d_intensity, d_severity, lowest_spei, month_peak, start_date, end_date, middate) |> 
  separate(sp_elev, c("sp_code", "elev_code"), sep = "_", remove = FALSE)

d12 |> 
  mutate(elev_code = fct_relevel(elev_code, "low-Dec", "low", "medium", "high")) |>
  mutate(elev_code = fct_relevel(elev_code, "low-Dec", "low", "medium", "high")) |>
  ggplot(aes(y = elev_code)) +
  geom_point(aes(x = middate, y = elev_code, size = d_severity, colour = sp_code)) + 
  geom_segment(aes(x = start_date, xend = end_date, yend = elev_code), 
               colour = "black", lineend = "butt", linewidth = 1.5) + 
  labs(size = "Drougth's Severity", x="", y = "Elevation") + 
  # facet_wrap(.~sp_code, ncol = 1, scales = "free_y") +
  facet_wrap(~factor(sp_code, levels = c("sylvestris", "nigra", "pinaster","halepensis")), ncol = 1, scales = "free_y") +
  theme_bw() +
    scale_colour_manual(values = colours_sp, name = "Species") + 
  scale_x_date(breaks = seq.Date(from = ymd("1980-01-01"), to = ymd("2022-01-01"), by = "5 years"),
               date_labels = "%Y",
               date_minor_breaks = "1 year") +
  theme(strip.background = element_blank())

```

```{r}
d12 |> 
  kbl(caption = "Drought periods by site (SPEI-12)") |> 
    kable_styling()
```


## Summary drought events

```{r}
events_sp_elev <- d12 |> 
  group_by(sp_elev,sp_code) |> 
  count()


events_sp_elev |> 
  group_by(sp_code) |> 
  summarise(mean= mean(n))

```



```{r}
events_by_site <- d12 |> 
  group_by(
    sp_code, sp_elev, elev_code) |> 
  summarise(
    across(c(starts_with("d"), "lowest_spei"),
           list(mean = mean, sd = sd, n = length), .names = "{.col}.{.fn}"))
```


```{r}
