---
title: "Modelling Ratio Selected ABI:NPP"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
--- 

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```


```{r, message=FALSE}
library(tidyverse)
library(patchwork)
library(likelihood)
library(kableExtra)
library(patchwork)
library(ggnewscale)
source("scripts/aux.R")
library(sf)
library(ggh4x)
library(geomtextpath)
library(metR)
library(ggsci)
library(fields)
library(splines)

```

```{r}
annual_pet <- read_csv("data/dendroadaptamed_spei_climate.csv") |> 
  dplyr::select(sp_elev, year, monthly_pet, monthly_tmed, monthly_prec) |> 
  group_by(sp_elev, year) |>
  summarise(pet = sum(monthly_pet), 
            prec = sum(monthly_prec),
            tmed = mean(monthly_tmed, na.rm = TRUE)) |> 
  rowwise() |> 
  mutate(water_balance = prec - pet)
```


```{r}
ratio <- read_csv("data/dendroadaptamed_ratio_abinpp.csv") |> 
  dplyr::select(year, sp_code, elev_code, sp_elev, ratio_abi_npp = ratio) 

df <- ratio |> 
  inner_join(annual_pet) |> 
  pivot_longer(pet:water_balance, values_to = "mean_climate", 
               names_to = "climate_variable")

d <- df |> filter(climate_variable %in% c("tmed", "prec")) |> 
  pivot_wider(values_from = mean_climate, names_from = climate_variable) |> 
  dplyr::rename(rat = ratio_abi_npp) |> 
  mutate(species = paste0("P. ", sp_code))
```

# Temperature 
- Tmed: Log-normal and Log-modified

$$(ABI:NPP)_i=MR \times g(T_{med}) + \epsilon_i$$

- Log-normal (**mtrat3**)
$$g(T_{med})=\exp \left[ -\frac{1}{2} \left(\frac{\log{(\frac{T_{med}}{OT_{med}}}) }{b}\right)^2 \right]$$ 
siendo $OT_{med}$ el valor óptimo de $T_{med}$ al cual ocurre el máximo Ratio $(ABI:NPP)_i$; y $b$ es la desviación estándar o amplitud de la función. 

- Logistic modified (**mtrat6*)

$$g(T_{med})=\left(1 - b \right) \times \left[ \frac{1}{1+\exp \left(-k \times (T_{med} - t0) \right) } \right] + b$$ 




```{r}
var=list(mean = "predicted", x = "rat", tmed = "tmed", prec = "prec", log = TRUE)
```

```{r}
# temp Log-normal
set.seed(123456)

mtrat3 <- function(MR, OTmed, Tb)
{
  tmed_effect <- exp(-0.5*(log(d$tmed/OTmed)/Tb)^2)
  MR * tmed_effect
}

# change model 
result_mtrat3 <- anneal(
  model = mtrat3, 
  var = var, 
  source_data = d,
  par = list(MR = 0.7, OTmed = 10.5, Tb = 1.5),
  par_lo = list(MR = 0, OTmed = 0, Tb = 0.1),
  par_hi = list(MR = 1.2, OTmed = 20, Tb = 5),
  pdf = dnorm,
  dep_var = "rat",
  initial_temp = 5, 
  temp_red = 0.975, 
  max_iter = 300000, 
  show_display = FALSE
)

result_mtrat3 |> likelihood::write_results("output/models/univariate_temp_lognormal.txt")
```



```{r}
# temp Logistic_mod
set.seed(1234)

mtrat6 <- function(MR,k,t0, b){
  tmed_effect <- (1-b)*(1 / (1 + exp(- k *( d$tmed - t0)))) + b 
  MR*tmed_effect
}


result_mtrat6 <- anneal(
  model = mtrat6, 
  var = var, 
  source_data = d,
  par = list(MR = 0.7, b = 0.09, k = -1, t0 = 10),
  par_lo = list(MR = 0, b = 0, k = -2.5, t0 = 0),
  par_hi = list(MR = 1.2, b = 0.2, k = 2, t0 = 20),
  pdf = dnorm,
  dep_var = "rat",
  initial_temp = 5, 
  temp_red = 0.975, 
  max_iter = 200000, 
  show_display = FALSE
)

result_mtrat6 |> likelihood::write_results("output/models/univariate_temp_logisticmod.txt")
```


# Prec 
- Precipitation: Log-normal 



$$(ABI:NPP)_i=MR \times g(P) + \epsilon_i$$
- Log-normal (**mprecrat3**)
$$g(P)=\exp \left[ -\frac{1}{2} \left(\frac{\log{(\frac{P}{P{opt}}}) }{Pb}\right)^2 \right]$$ 

```{r}
## Log-normal 
set.seed(1234)

mprecrat3 <- function(MR, Popt, Pb)
{
  prec_effect <- exp(-0.5*(log(d$prec/Popt)/Pb)^2)
  MR * prec_effect 
}


# change model 
result_mprecrat3 <- anneal(
  model = mprecrat3, 
  var = var, 
  source_data = d,
  par = list(MR = 0.7, Popt = 500, Pb = 2),
  par_lo = list(MR = 0, Popt = 300, Pb = 0),
  par_hi = list(MR = 1, Popt = 1000, Pb = 5),
  pdf = dnorm,
  dep_var = "rat",
  initial_temp = 5, 
  temp_red = 0.975, 
  max_iter = 200000, 
  show_display = FALSE
)

result_mprecrat3 |> likelihood::write_results("output/models/univariate_prec_lognormal.txt")
```


## Multiplicative 

- ***Model combrat1***
    - log-normal temperature
    - log-normal prec

```{r}
mcombrat1 <- function(MR, OTmed, Tb, Popt, Pb){
  # log-normal temperature
  # log-normal prec 
  tmed_effect <- exp(-0.5*(log(d$tmed/OTmed)/Tb)^2) 
  prec_effect <- exp(-0.5*(log(d$prec/Popt)/Pb)^2)
  MR * tmed_effect * prec_effect
}

set.seed(2023)

# change model 
result_mcombrat1 <- anneal(
  model = mcombrat1, 
  var = var, 
  source_data = d,
  par = list(MR = 0.7, OTmed = 10.5, Tb = 1.5, Popt = 500, Pb = 2), 
  par_lo = list(MR = 0, OTmed = 5, Tb = 0.01, Popt = 300, Pb = 0),
  par_hi = list(MR = 1, OTmed = 20, Tb = 5, Popt = 1000, Pb = 5),
  pdf = dnorm,
  dep_var = "rat",
  initial_temp = 5, 
  temp_red = 0.975, 
  max_iter = 200000, 
  show_display = FALSE
)

result_mcombrat1 |> likelihood::write_results("output/models/multiplicative_tmed_lognormal_prec_lognormal.txt")
```


- ***Model combrat7***
    - log-mod  temperature
    - log-normal prec

```{r}
mcombrat7 <- function(MR, Tb, k, Topt, Popt, Pb){
  # log-mod temperature
  # log-normal prec 
  tmed_effect <- (1-Tb)*(1 / (1 + exp(- k *( d$tmed - Topt)))) + Tb
  prec_effect <- exp(-0.5*(log(d$prec/Popt)/Pb)^2)
  MR * tmed_effect * prec_effect
}

set.seed(123456)

result_mcombrat7 <- anneal(
  model = mcombrat7, 
  var = var, 
  source_data = d,
  par = list(MR = 0.7, Tb = 0.5, k = -1, Topt = 10, Popt = 500, Pb = 2), 
  par_lo = list(MR = 0, Tb = 0, k = -2, Topt = 0, Popt = 300, Pb = 0),
  par_hi = list(MR = 1, Tb = 1, k = 2, Topt = 20, Popt = 1000, Pb = 5),
  pdf = dnorm,
  dep_var = "rat",
  initial_temp = 5, 
  temp_red = 0.975, 
  max_iter = 200000, 
  show_display = FALSE
)


result_mcombrat7 |> likelihood::write_results("output/models/multiplicative_tmed_logisticmod_prec_lognormal.txt")
```


# Compare models 

```{r}
name_models <- data.frame(
  modelo = c(paste0("mtrat",c(3,6)), "mprecrat3", "mcombrat1", "mcombrat7"), 
  name_modelo = c("Temp (Log-normal)", "Temp (Logistic_Mod)", "Prec (Log-normal)",
                  "Temp (Log-normal) | Prec (Log-normal)", 
                  "Temp (Logistic_Mod) | Prec (Log-normal)"), 
  type = c(rep("univariate",3), rep("multiplicative",2))
)

all_results <-
  mget(paste0("result_", name_models$modelo)) |>
  purrr::imap_dfr(~ MLE_results_format(.x, yvar = "rat"), .id = "modelo") |>
  arrange(aic_cor) |>
  mutate(modelo = str_remove(modelo, "result_")) |> 
  inner_join(name_models) |>
  relocate(name_modelo, type, .after = modelo) 
```

```{r}
all_results |> 
  mutate(deltaAIC = dAIC(aic_cor)) |> 
  mutate(w = wAIC(deltaAIC)) |> 
  kbl(digits = c(0,0,0,2,0,2,2,3,3,2,3,3)) |> 
  kable_styling()
```


```{r}
univariate <- all_results |> 
  filter(type == "univariate") |> 
  mutate(deltaAIC = dAIC(aic_cor)) |> 
  mutate(w = wAIC(deltaAIC)) 

univariate |> 
  kbl(digits = c(0,0,0,2,0,2,2,3,3,2,3,3)) |> 
  kable_styling()
```


```{r}
multiplicative  <- all_results |> 
  filter(type == "multiplicative") |> 
  mutate(deltaAIC = dAIC(aic_cor)) |> 
  mutate(w = wAIC(deltaAIC))

multiplicative |> 
  kbl(digits = c(0,0,0,2,0,2,2,3,3,2,3,3)) |> 
  kable_styling()
```

<!-- 
Cuando el likelihood es similar, aplicar el Ockham’s razor (mas parsimonioso)

- Usar el modelo con menor valor de AIC (mas cercano a la "verdad")
- Ojo el AIC, identifica el mejor modelo de un conjunto, aún cuando los modelos sean pobres 
- Los modelos incluidos en el conjunto de modelos, tienen que ser modelos realistas, con sentido ecológico (no cualquier modelo) 
- Usar AICc si n/k (size / nparams) es < 40 (ojo usar siempre o AIC o AICc, pero no mezclar)
- BIC 

- Differences in AIC (Δi’s) can be used to interpret strength of evidence for one model vs. another.
- A model with a Δ value within 1-2 of the best model has substantial support in the data, and should be considered along with the best model.
- A Δ value within only 4-7 units of the best model has considerably less support.
- A Δ value > 10 indicates that the worse model has virtually no support and can be omitted from further consideration.

- Akaike weights 
--> 

### Logistic mod & Log_normal 

```{r}
colours_combined_name <- c(
                      "Temp (Log-normal)" = "red", 
                      "Temp (Logistic_Mod)" = "black",
                      "Prec (Log-normal)" = "black")

colours_combined <- c(
                      "mtrat3" = "red", 
                      "mtrat6" = "black",
                      "mprecrat3" = "black") 

alpha_points <- 0.5 

label_npp <- "Annual~NPP[MODIS]~(g~C~m^2~year^{-1})"
label_ratio <- "ABI:NPP"
label_prec <- "Precipitation (mm)"
label_tmed <- "Annual Mean Temperature (ºC)"
label_wb <- "Water Balance (mm)"
```


```{r}
ptlognormal  <- result_mtrat3$best_pars |> as.data.frame()
ptlogmod <- result_mtrat6$best_pars |> as.data.frame()
pplognormal <- result_mprecrat3$best_pars |> as.data.frame()


teffect <- data.frame(tmed = seq(5,25, by =.1)) |> 
  mutate(
    mtrat3_teffect = exp(-0.5*(log(tmed / ptlognormal$OTmed)/ptlognormal$Tb)^2),
    mtrat6_teffect = (1-ptlogmod$b)*(1 / (1 + exp(- ptlogmod$k *(tmed - ptlogmod$t0)))) + ptlogmod$b,
    mtrat3_predicted = ptlognormal$MR * mtrat3_teffect,
    mtrat6_predicted = ptlogmod$MR * mtrat6_teffect
  ) |> 
  pivot_longer(-tmed) |> 
  separate(name, into = c("modelo", "type_value")) |> 
  inner_join(name_models)

plot_tmed <- teffect |> 
  filter(type_value == "predicted") |> 
  ggplot(
    aes(x = tmed, y = value)) + 
  geom_line(linetype = "dashed", aes(colour = name_modelo)) + 
  geom_line(
    data = (teffect |>
              filter(type_value == "predicted") |> 
              filter(tmed > min(d$tmed)) |> 
              filter(tmed < max(d$tmed))), 
    aes(colour = name_modelo), 
    linewidth = 1.5) + 
  scale_colour_manual(values = colours_combined_name, name = "Models") +
  ylim(0, 0.7) +
  xlim(7.5, 17.5) + 
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab(label_tmed) +
  ylab(label_ratio) + 
  new_scale_colour() +
  geom_point(data = d, 
             aes(x = tmed, 
                 y = rat, 
                 shape = elev_code, fill = species, colour = species), 
             alpha = alpha_points) +
  scale_shape_manual(values = shape_elev, name = "Elevation") + 
  scale_colour_manual(
    values = colours_Specie, 
    labels = expression(italic("P. halepensis"), italic("P. pinaster"), italic("P. nigra"), italic("P. sylvestris")),
    name = "Species") + 
  scale_fill_manual(
    values = colours_Specie, 
    labels = expression(italic("P. halepensis"), italic("P. pinaster"), italic("P. nigra"), italic("P. sylvestris")),
    name = "Species")
```

```{r}
peffect <- data.frame(prec = seq(0,1700, by =1)) |> 
  mutate(
    mprecrat3_peffect = exp(-0.5*(log(prec/ pplognormal$Popt)/pplognormal$Pb)^2),
    mprecrat3_predicted = pplognormal$MR *mprecrat3_peffect) |> 
  pivot_longer(-prec) |> 
  separate(name, into = c("modelo", "type_value")) |> 
  inner_join(name_models)

plot_prec <- peffect |> 
  filter(type_value == "predicted") |> 
  ggplot(
    aes(x = prec, y = value)) + 
  geom_line(linetype = "dashed", aes(colour = name_modelo)) + 
  geom_line(
    data = (peffect |>
              filter(type_value == "predicted") |> 
              filter(prec > min(d$prec)) |> 
              filter(prec < max(d$prec))), 
    aes(colour = name_modelo), 
    linewidth = 1.5) + 
  scale_colour_manual(values = colours_combined_name, name = "Models") +
  ylim(0, 0.7) +
  # xlim(7.5, 17.5) + 
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab(label_prec) +
  ylab(label_ratio) + 
  new_scale_colour() +
  geom_point(data = d, 
             aes(x = prec, 
                 y = rat, 
                 shape = elev_code, fill = species, colour = species), 
             alpha = alpha_points) +
  scale_shape_manual(values = shape_elev, name = "Elevation") + 
  scale_colour_manual(
    values = colours_Specie, 
    labels = expression(italic("P. halepensis"), italic("P. pinaster"), italic("P. nigra"), italic("P. sylvestris")),
    name = "Species") + 
  scale_fill_manual(
    values = colours_Specie, 
    labels = expression(italic("P. halepensis"), italic("P. pinaster"), italic("P. nigra"), italic("P. sylvestris")),
    name = "Species")
```


```{r}
plot_tmed + plot_prec & theme(legend.position = "bottom") + plot_layout(guides = "collect")
```



# Predicted 

```{r}
pcomb1 <- result_mcombrat1$best_pars |> as.data.frame()
pcomb7 <- result_mcombrat7$best_pars |> as.data.frame()


pred_mcombrat7<- expand.grid(tmed = seq(5,20, by =.5),
                    prec = seq(100, 1500, by = 5)) |> 
  mutate(
    ratio_pred =   pcomb7$MR * ((1 - pcomb7$Tb)*(1 / (1 + exp(- pcomb7$k *(tmed - pcomb7$Topt)))) + pcomb7$Tb) * (exp(-0.5*(log(prec/pcomb7$Popt)/pcomb7$Pb)^2) )
  ) 

breaks <- seq(0, .75, by=0.05)


## plot surface  
plot_surface_mcombrat7 <- pred_mcombrat7 |> 
  ggplot(aes(x = prec, y = tmed)) +
  metR::geom_contour_fill(aes(z = ratio_pred), 
                          show.legend = FALSE, breaks = breaks, colour = "white") + 
  geom_point(
    data = d,
    colour = "white",
    size = 1.2
  ) + 
  scale_fill_material("blue-grey") + 
  metR::geom_label_contour(aes(z = ratio_pred),
                          breaks = breaks,  colour = "black", 
                          label.placer = label_placer_n(n=1))  +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Precipitation (mm)") +
  ylab("Annual Mean Temperature (ºC)") + 
  scale_y_continuous(
    limits = c(7.5, 17.5), 
    breaks = c(7.5, 10, 12.5, 15, 17.5)) + 
  scale_x_continuous(limits = c(50, 1500), 
                     breaks = c(150, 300, 450, 600, 750, 900, 1050, 1200, 1350, 1500))
```



```{r}
a <- plot_tmed + plot_prec & theme(legend.position = "bottom") + plot_layout(guides = "collect")

combined7 <- a + plot_surface_mcombrat7
```


```{r}
pcomb1 <- result_mcombrat1$best_pars |> as.data.frame()

pred_mcombrat1<- expand.grid(tmed = seq(5,20, by =.5),
                    prec = seq(100, 1500, by = 5)) |> 
  mutate(
    ratio_pred =   pcomb1$MR * (  exp(-0.5*(log(tmed/pcomb1$OTmed)/pcomb1$Tb)^2))* (exp(-0.5*(log(prec/pcomb1$Popt)/pcomb1$Pb)^2) )
  ) 


breaks <- seq(0, 1, by=0.05)


## plot surface  
plot_surface_mcombrat1 <- pred_mcombrat1 |> 
  ggplot(aes(x = prec, y = tmed)) +
  metR::geom_contour_fill(aes(z = ratio_pred), 
                          show.legend = FALSE, breaks = breaks, colour = "white") +
  geom_point(
    data = d,
    colour = "white",
    size = 1.2
  ) + 
  scale_fill_material("blue-grey") + 
  metR::geom_label_contour(aes(z = ratio_pred),
                          breaks = breaks,  colour = "black", 
                          label.placer = label_placer_n(n=1))  +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Precipitation (mm)") +
  ylab("Annual Mean Temperature (ºC)") + 
  scale_y_continuous(
    limits = c(7.5, 17.5), 
    breaks = c(7.5, 10, 12.5, 15, 17.5)) + 
  scale_x_continuous(limits = c(50, 1500), 
                     breaks = c(150, 300, 450, 600, 750, 900, 1050, 1200, 1350, 1500))
```

```{r}
a <- plot_tmed + plot_prec & theme(legend.position = "bottom") + plot_layout(guides = "collect")

combined1 <- a + plot_surface_mcombrat1
```

## Main figure 

```{r}
line_colour_model <- "black"


custom_options <- list(
  scale_shape_manual(
    values = shape_elev, 
    name = "Elevation"), 
  scale_colour_manual(
    values = colours_Specie, 
    labels = expression(italic("P. halepensis"), italic("P. pinaster"), italic("P. nigra"), italic("P. sylvestris")),
    name = "Species"),
  scale_fill_manual(
    values = colours_Specie, 
    labels = expression(italic("P. halepensis"), italic("P. pinaster"), italic("P. nigra"), italic("P. sylvestris")),
    name = "Species"),
  theme_bw(),
  theme(
    panel.grid = element_blank(), 
    axis.title = element_text(size = 14, face = "bold"), 
    axis.text = element_text(size = 11),
    axis.ticks.length = unit(.2, "cm"),
    legend.title=element_text(size=12, face = "bold"), 
    legend.text=element_text(size=12)
    )
)

# plot temp 
plot_tmed_m7 <- 
  teffect |>
  filter(type_value == "predicted") |> 
  filter(modelo == "mtrat6") |> 
  ggplot(
    aes(x = tmed, y = value)) + 
  geom_line(
    linetype = "dashed",
    colour = line_colour_model) + 
  geom_point(data = d, 
             aes(x = tmed, 
                 y = rat, 
                 shape = elev_code, fill = species, colour = species), 
             alpha = alpha_points) +
  geom_line(
    data = (teffect |>
              filter(type_value == "predicted") |> 
              filter(modelo == "mtrat6") |> 
              filter(tmed > min(d$tmed)) |> 
              filter(tmed < max(d$tmed))),
    colour = line_colour_model, 
    size = 1.5) +
  scale_y_continuous(limits = c(0,0.63), expand = expansion(add = c(0.005, 0.005))) + 
  scale_x_continuous(limits = c(7.5,17.8), 
                     expand = expansion(add = c(0.1, 0.1)),
                     breaks = seq(7.5, 17.5, 2.5)) + 
  xlab(label_tmed) +
  ylab(label_ratio) +
  custom_options + 
  annotate(geom="text", 
           x=15, y=.6, 
           label= sprintf("R^2 == %.3g",round(result_mtrat6$R2,3)), 
           color="black",
           parse = TRUE, 
           size = 6)
  


plot_prec_m7 <-
  peffect |> 
  filter(type_value == "predicted") |> 
  ggplot(
    aes(x = prec, y = value)) + 
  geom_line(
    linetype = "dashed",
    colour = line_colour_model) + 
  geom_point(data = d, 
             aes(x = prec, 
                 y = rat, 
                 shape = elev_code, fill = species, colour = species), 
             alpha = alpha_points) +
  geom_line(
    data = (peffect |>
              filter(type_value == "predicted") |> 
              filter(prec > min(d$prec)) |> 
              filter(prec < max(d$prec))), 
    colour = line_colour_model, 
    size = 1.5) + 
  scale_y_continuous(limits = c(0,0.63), expand = expansion(add = c(0.005, 0.005))) + 
  scale_x_continuous(limits = c(50, 1550), 
                     breaks = c(150, 300, 450, 600, 750, 900, 1050, 1200, 1350, 1500)) + 
  xlab(label_prec) + 
  ylab("") + 
  custom_options +
  annotate(geom="text", 
           x=1050, y=.6, 
           label= sprintf("R^2 == %.3g",round(result_mprecrat3$R2,3)), 
           color="black",
           parse = TRUE, 
           size = 6) 


breaks <- seq(0, .8, by=0.05)

# Generate clip 
x <- d[chull(d$tmed, d$prec), c("rat","prec","tmed")] |> 
  add_row(rat = 0.075, prec = 800, tmed = 12.5) |> 
  mutate(prec = ifelse(prec < 150, prec - 30, 
                        ifelse(prec > 1300, prec + 30, prec)))

bb <- st_as_sf(x = x, 
               coords = c("prec","tmed"), remove = FALSE) |> 
  rename(ratio_pred = rat) |> 
  st_combine() |>   # combine to generate polygon
  st_cast("POLYGON") |> 
  st_buffer(dist = 1)

plot_surface_mcombrat7 <- pred_mcombrat7 |> 
  ggplot(aes(x = prec, y = tmed)) +
  metR::geom_contour_fill(aes(z = ratio_pred, 
                              fill = after_stat(level)), 
                          binwidth = 0.1,
                          show.legend = TRUE, 
                          breaks = breaks, 
                          colour = "white", clip = bb) + 
  geom_point(
    data = d,
    colour = "black", alpha = 0.4, 
    size = 1.2
  ) + 
  metR::geom_label_contour(aes(z = ratio_pred),
                          colour = "black", 
                          breaks = breaks, 
                          label.placer = label_placer_n(n=1))  +
  scale_fill_discretised(name = label_ratio, 
                         low = "#ffffe5", high = "#004529", 
                         guide = guide_legend(theme = theme(
                           legend.title=element_text(size=11, face = "bold"),
                           legend.text=element_text(size=9)), 
                           ncol = 2
                         )) + 
  theme_bw() +
  theme(
    panel.grid = element_blank(), 
    axis.title = element_text(size = 14, face = "bold"), 
    axis.text = element_text(size = 11),
    axis.ticks.length = unit(.2, "cm"),
    legend.position = c(0.75,0.78)) + 
  xlab(label_prec) + 
  ylab(label_tmed) + 
  scale_y_continuous(
    limits = c(7.5, 17.5), 
    breaks = c(7.5, 10, 12.5, 15, 17.5),
    expand = expansion(add = c(0.1, 0.01))) + 
  scale_x_continuous(
    limits = c(50, 1510), 
    breaks = c(150, 300, 450, 600, 750, 900, 1050, 1200, 1350, 1500),
    expand = expansion(add = c(1, 1))) 

```

```{r}
main_figure <- (
  (plot_tmed_m7 + plot_prec_m7 & theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.box.just = "left"
  )) + plot_layout(guides = "collect")
    + (plot_surface_mcombrat7 + plot_layout(guides = "keep"))) +
  plot_annotation(tag_levels = "A") & theme(
    plot.tag = element_text(size = 14, face = "bold"))


ggsave(
   main_figure,
   file = "output/figure_model_selected.jpg",
   dpi = 600,
   width = 15, height = 7
 )
```


```{r}
#| fig.height: 7
#| fig.width: 15
main_figure
```


### Model validation
```{r}
#| echo: false
#| fig.show: hide

plot_observed <- function(x, yvar, annotate = TRUE, 
                          lab_x = "Observed", 
                          lab_y = "Predicted", ...) { 
  
  d <- x$source_data |> 
    mutate(residuals = !!sym(yvar)  - predicted) |> 
    rename(observed = !!sym(yvar))
  
  model_results <- MLE_results_format(x, yvar = "rat")
  
  max_value <- max(max(d$predicted, na.rm = TRUE), max(d$observed, na.rm = TRUE))
  max_range <- max(0, max_value)
  
  p <- ggplot(d, aes(x = predicted, y = observed)) +
    labs(x = lab_x, y = lab_y) + 
    geom_point() +
    geom_abline() +
    theme_bw() + 
    theme(panel.grid = element_blank()) +
    xlim(0, max_range) + 
    ylim(0, max_range)
  
  if (annotate) { 
    out_p <- p + 
      annotate("text", 
               x = 0.01 * max_range, y = 0.9 * max_range, 
               label = paste("R^2~'='", sprintf("%.3f", model_results$R2), sep="~"), 
               hjust = 0, vjust = 0, parse = TRUE) +
      annotate("text", 
               x = 0.01 * max_range, y = 0.85 * max_range, 
              label = paste("Slope~'='", sprintf("%.4f", model_results$slope), sep="~"), 
               hjust = 0, vjust = 0, parse = TRUE)
  } else {
    out_p <- p 
  }
  
  print(out_p)
}
  
  
p_observed <- plot_observed(result_mcombrat7, yvar = "rat", 
                            lab_y = "Observed ABI:NPP", 
                            lab_x= "Predicted ABI:NPP") 
```

```{r}
#| echo: false
#| fig.show: hide
plot_residuals <- function(x, yvar, lab_residuals = "Residuals", 
                           lab_predicted = "Predicted", ...) { 
  
  d <- x$source_data |> 
    mutate(residuals = !!sym(yvar)  - predicted) |> 
    rename(observed = !!sym(yvar))
  
  
  out <- ggplot(d, aes(x = predicted, y = residuals)) +
    labs(x = lab_predicted, y = lab_residuals) + 
    geom_point() +
    geom_hline(yintercept = 0) +  
    xlim(0, NA) +
    theme_bw() + 
    theme(panel.grid = element_blank()) 
  
  print(out)
}


p_residuos <- plot_residuals(result_mcombrat7, yvar = "rat", lab_residuals = "Residuals ABI:NPP", 
               lab_predicted = "Predicted ABI:NPP") 

```


```{r}
#| echo: false
prepare_data <- function(x, yvar) {
  x$source_data |> 
    mutate(residuals = !!sym(yvar)  - predicted) |> 
    rename(observed = !!sym(yvar)) 
}

h <- prepare_data(result_mcombrat7, yvar = "rat")

dist_residuos <- ggplot(h, aes(x = residuals)) +
  geom_histogram(aes(y = after_stat(density), x = residuals), fill = "gray", col = "white") +
  stat_function(fun = dnorm, 
                args = list(mean = mean(h$residuals, na.rm = TRUE), 
                            sd = sd(h$residuals, na.rm = TRUE)), 
                color = "black", size = .85) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Residuals") +
  ylab("Density")
```


```{r}
plot_validation <- (p_observed + p_residuos + dist_residuos +  plot_annotation(tag_levels = "A")) & 
  theme(axis.title = element_text(size = 14, face = "bold"), 
        axis.text = element_text(size = 11),
        plot.tag = element_text(size = 14, face = "bold"))

ggsave(
   plot_validation,
   file = "output/figure_model_validation.jpg",
   dpi = 600,
   width = 13, height = 5.5
 )
```

```{r}
#| fig.height: 5
#| fig.width: 12
plot_validation
```



### Tables 
- support limits 
```{r}
support_limits_table <- function(x) { 
  
  data <- x
  out <- inner_join(
  (data$lower_limits |> bind_rows() |> pivot_longer(everything(), names_to = "parameter", values_to = "lower_si")),
  (data$upper_limits |> bind_rows() |> pivot_longer(everything(), names_to = "parameter", values_to = "upper_si"))
)
  return(out)
}


set.seed(123456)
limits <- support_limits(mcombrat7, result_mcombrat7$best_pars, var = var, 
                         source_data = d, pdf = dnorm,
                         par_lo = list(MR = 0, Tb = -1, k = 1, Topt = , Popt = 50, Pb = 0),
                         par_hi = list(MR = 1, Tb = 1, k = 2, Topt = 20, Popt = 3500, Pb = 2),
                         slimit = 2)


table_support_limits <- inner_join(
  result_mcombrat7$best_pars |> bind_rows() |> pivot_longer(everything(), names_to = "parameter", values_to = "best_pars"),
           support_limits_table(limits))

table_support_limits

table_support_limits |> write_csv("output/models/table_params_support_limit.csv")
```


```{r}
#| eval: false
#| echo: false

# He generado un algoritmo de fuerza bruta para que me busque los SI de los parametros
# que no alcanzan un valor (lower or upper limit = par_lo or par_hi)

set.seed(123456)

out <- c()
for (j in seq(-50, 30, by = 0.5)){ 
  
  limits <- support_limits(mcombrat7, result_mcombrat7$best_pars, var = var, 
                         source_data = d, pdf = dnorm,
                         par_lo = list(MR = 0, Tb = -1, k = 1, Topt = j, Popt = 50, Pb = 0),
                         par_hi = list(MR = 1, Tb = 1, k = 2, Topt = j+10, Popt = 3500, Pb = 2),
                         slimit = 2)
  
  limits_out <- support_limits_table(limits) |> mutate(j_value = j)
  
  out <- rbind(out, limits_out)
  }

o <- out |> filter(parameter == "Topt") |> 
  pivot_longer(c(lower_si, upper_si))

o |> ggplot(aes(x = j_value, y = value, colour=name)) + geom_point()


```

