---
title: "Temporal evolution of NPP from MODIS MOD17A3"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set( 
                      warning=FALSE, message=FALSE)
```

# Aims

- Prepare temporal series of yearly NPP for each field-plot, and then compute temporal trends

- Data source: 
  
  - MODIS [MOD17A3HGF](https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MOD17A3HGF)
  - Spatial resolution: 500 m pixel
  - The MOD17A3HGF V6.1 product provides information about annual Gross and Net Primary Productivity (GPP and NPP) at 500m pixel resolution. Annual NPP is derived from the sum of all 8-day Net Photosynthesis(PSN) products (MOD17A2H) from the given year. The PSN value is the difference of the Gross Primary Productivity (GPP) and the Maintenance Respiration (MR) (GPP-MR).
  - Detailed info [here](https://lpdaac.usgs.gov/documents/972/MOD17_User_Guide_V61.pdf)

```{r pkg}
library(here)
library(tidyverse)
library(tidytext)
library(Kendall)
library(trend)
library(kableExtra)
library(DT)

source("scripts/aux.R")
```

```{r}
dnpp <- read_csv("data/dendroadaptamed_npp_mod17a3_yearly_c61.csv") |> filter(year <= 2022)
```


```{r}
npp_summary <- dnpp |> 
  mutate(elev_code = fct_relevel(elev_code, c("high","medium","low", "low-Dec"))) |> 
  group_by(Specie, elev_code, sp_elev) |> 
  summarise(mean = mean(npp, na.rm = TRUE), 
            sd = sd(npp, na.rm = TRUE), 
            se = sd/sqrt(length(npp)),
            cv = sd/mean*100,
            min = min(npp), 
            max = max(npp)) 

```

```{r npp-summary}
npp_summary |> 
  dplyr::select(-sp_elev) |> 
  arrange(Specie, elev_code) |> 
  kbl(
    digits = c(0,0,0,2,2,2,2,2,2),
    caption = "Summary values of the yearly NPP (g C m^-2^ year ^-1^) along the study period (2000-2021) for the study sites.", parse = TRUE) |> 
    column_spec(1, italic = TRUE) |> 
  collapse_rows(columns = 1:2) |> 
    kable_paper("hover", full_width = F)
```


```{r, plot-npp-avg}
#| fig.cap: "Comparison of average values of yearly NPP."
#| fig.width: 12
#| fig.height: 8

npp_summary |> 
  mutate(elev_code = fct_relevel(elev_code, c("low-Dec","low","medium","high"))) |> 
  ggplot(aes(x=elev_code, y=mean, fill=elev_code, colour=elev_code)) + 
  geom_point(size=3) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.3)+
  ylab(expression(NPP~(g~C~m^{-2}~year^{-1}))) + 
  xlab("") + 
  facet_grid(cols=vars(factor(Specie, levels = c("P. halepensis", "P. pinaster", "P. nigra", "P. sylvestris"))), scales = "free_x", space = "free_x") +
  scale_fill_manual(values=colours_elev, name="") +
  scale_colour_manual(values=colours_elev, name="") + 
  theme_bw() + 
    theme(
    panel.grid = element_blank(),
    # panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill=NA), 
    strip.text = element_text(face="italic")
  )

```


# Compute Temporal trends

We explored the temporal trends for the NPP annual average (Table \@ref(tab:mk-mean)) along study-period. For this purpose, we use non-parametric Mann-Kendall test. We also explored the magnitude of the trends using the Theil-Sen's slope estimator. MK was computed from 2001 to 2022 of MODIS NPP.

```{r}
mknpp <- dnpp |> 
  ungroup() |> 
  group_by(Specie, elev_code, sp_elev, sp_code) |> 
  summarise(across(npp, ~MannKendall(.)$tau, .names ="{.col}_tau"),
            across(npp, ~MannKendall(.)$sl, .names ="{.col}_pvalue_mk"),
            across(npp, ~trend::sens.slope(.)$estimate, .names ="{.col}_senslope"),
            across(npp, ~trend::sens.slope(.)$p.value, .names ="{.col}_pvalue_sen")) |> 
  arrange(Specie, elev_code) |> 
  mutate(ypos = 
           case_when(
             elev_code == "low" ~ 1025,
             elev_code == "low-Dec" ~ 1000,
             elev_code == "medium" ~ 1050,
             elev_code == "high" ~ 1075), 
         p_value_string = symnum(npp_pvalue_mk, corr = FALSE,
               cutpoints = c(0,  .001,.01,.05, 1),
               symbols = c("***","**","*",""))) |> 
  mutate(taulabel = paste(expression(tau), "==", paste0('"', round(npp_tau, 2), p_value_string, '"')))

```


```{r mk-mean}
mknpp |> 
  ungroup() |> 
  dplyr::select(-ypos, -p_value_string, -taulabel, -sp_code, -sp_elev) |> 
  kbl(
    digits = c(0,0,3,4,5,4),
    caption = "Mann-Kendall and Seil temporal trend test for yearly NPP") |> 
    column_spec(3.4, bold = ifelse(mknpp$npp_pvalue_mk < 0.05, TRUE, FALSE)) |> 
  column_spec(1, italic = TRUE) |> 
  collapse_rows(columns = 1:2) |> 
    kable_paper("hover", full_width = F)
```


```{r plot-npp}
#| fig.cap: "Temporal evolution of year NPP."
#| fig.width: 12
#| fig.height: 8

dnpp |> 
  ggplot(aes(x=year, y=npp, colour=elev_code)) +
  geom_line() + 
  #geom_point() + 
  facet_wrap(~(factor(Specie, levels = c("P. halepensis", "P. pinaster", "P. nigra", "P. sylvestris"))),
             nrow = 1) +
  scale_fill_manual(values=colours_elev, name="") +
  scale_colour_manual(values=colours_elev, name="") +
  ylab(expression(NPP~(kg~C~m^{-2}~year^{-1}))) + 
  xlab("") + 
  theme_bw() + 
    theme(
    panel.grid = element_blank(),
    # panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill=NA), 
    strip.text = element_text(face="italic")
  ) +
  geom_text(x = 2002, aes(y = ypos, label = taulabel), data = mknpp, parse = TRUE, show.legend = FALSE, hjust = "left")
```


```{r}
write_csv(mknpp, "data/dendroadaptamed_npp_modis_trend.csv")
write_csv(dnpp, "data/dendroadaptamed_npp_modis.csv")
```


