---
title: "Scripts"
output: workflowr::wflow_html
bibliography: references.bib
csl: ecology-letters.csl
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set( 
                      warning=FALSE, message=FALSE)
```


```{r pkg}
library(here)
library(tidyverse)
```

# Scripts

This page contains some of the auxiliary and/or external scripts used in this project 

## Get Satellite data

Using Google Earth Engine, we extracted spectral information for all the pixels covering the field plots. Several products were obtained:

- **Temporal series of Vegetation Indexes** NDVI and EVI from MOD13Q1. (MODIS Collection 6.1) We used the following GEE code to get the EVI and NDVI data for each field plot. We obtained a total of 503 values for each field plot (2000-2021, 23 images per year, except 2000 year (only 20 images)). The data were saved as `mod13q1_iv_c61.csv` (see [/data/raw/remote_sensing/mod13q1_iv_c61.csv](https://github.com/ajpelu/adaptamed_dendroanalysis/blob/master/data/raw/remote_sensing/mod13q1_iv_c61.csv)). The temporal resolution was one image each 16 days (23 images per year), and the spatial resolution was 250 m.  


```{js, eval=FALSE}
var parcelas_old = ee.FeatureCollection("projects/ee-ajpelu/assets/geo_pinos_dendro_adaptamed");
var parcelas = ee.FeatureCollection("projects/ee-ajpelu/assets/geo_pinos_dendro_adaptamed_mod");

// Define study period 
var startdate = ee.Date('2000-01-01');
var enddate = ee.Date('2022-10-31'); 

//-------------------- Load MODIS MOD13Q1 
// filter by date 
// select bands: EVI, NDVI, DOY, SummaryQA
var mod13 = ee.ImageCollection('MODIS/061/MOD13Q1')
  .filterDate(startdate, enddate)
  .select(['EVI', 'NDVI']); 

print(mod13); 


// Apply a function to every image of the collection (use .map()) 
// Get value of each band b of each temporal image t of MODQ13Q1
// for each of the coordinate provide in parcelas
var mydata = mod13.map(function(i) {
  return i.reduceRegions(parcelas, 'first');
});

// TODO: Complete doc 
var extract = ee.FeatureCollection(mydata.flatten());

// Export table
Export.table.toDrive({
  collection: extract,
  description: 'my_extract_data',
  fileFormat: 'CSV',
});

```

- **Temporal series of NPP and GPP** from MOD17A3. We used the following GEE code to get the annual NPP and GPP data for each field plot. We obtained a total of 22 values for each field plot (2000-2021; 1 value per year). The data were saved as `dendroadaptamed_npp_mod17a3_yearly.csv` (see [data/dendroadaptamed_npp_mod17a3_yearly.csv](https://github.com/ajpelu/adaptamed_dendroanalysis/blob/master/data/dendroadaptamed_npp_mod17a3_yearly.csv)). Temporal resolution was yearly; Spatial resolution was 500 m. See [here](https://lpdaac.usgs.gov/products/mod17a3hgfv006/) for a description of the method to obtain the NPP. 

```{js, eval=FALSE}
var parcelas = ee.FeatureCollection("users/ajpelu/geoinfo_adaptamed_dendro");

// Define study period 
var startdate = ee.Date('2000-01-01');
var enddate = ee.Date('2023-10-31'); 

//-------------------- Load MODIS MOD17A3HGF.006 
// filter by date 
// select bands: Npp, Npp_QC
var modNpp = ee.ImageCollection('MODIS/061/MOD17A3HGF')
  .filterDate(startdate, enddate)
  .select(['Gpp','Npp', 'Npp_QC']); 

print(modNpp);


// Apply a function to every image of the collection (use .map()) 
// Get value of each band b of each temporal image t of MODIS
// for each of the coordinate provide in parcelas
var mydata = modNpp.map(function(i) {
  return i.reduceRegions(parcelas, 'first');
});

// TODO: Complete doc 
var extract = ee.FeatureCollection(mydata.flatten());

// Export table
Export.table.toDrive({
  collection: extract,
  description: 'my_extract_data',
  fileFormat: 'CSV',
});

```

- **Long-term series of NDVI**, using the *Sp_1km_NDVI* dataset. This dataset was derived from Atmospheric Administration–Advanced Very High Resolution Radiometer
(NOAA–AVHRR) by @VicenteSerranoetal2020VegetationGreening. The temporal resolution was bi-weekly (24 data by year) and the spatial resolution was 1.1 km. The data are available in https://spainndvi.csic.es/, but we could not extract the data for our study site locations. Nevertherless, we used the [raw data](https://spainndvi.csic.es/nc) and this [repository](https://github.com/ncwebmapper/ncwebmapper). Then we write a custom [script](https://github.com/ajpelu/adaptamed_dendroanalysis/blob/master/scripts/ndvi_csic.R) to downloaded the raw data in `.nc` format, prepare and extract the long-term series for our study sites. The data were saved as `dendroadaptamed_ndvicsic1km.csv` (see [data/dendroadaptamed_ndvicsic1km.csv](https://github.com/ajpelu/adaptamed_dendroanalysis/blob/master/data/dendroadaptamed_ndvicsic1km.csv)). 

## Auxiliary

- Categorize the elevation of each site
```{r}
elev_cat <- data.frame(site_code=
c("PIHAL_NACIMIENTO",
"PINIG_LANTEIRA_H","PINIG_LANTEIRA_L2","PINIG_LANTEIRA_M",
"PIPIN_JEREZ_H","PIPIN_JEREZ_L","PIPIN_JEREZ_M","PIPIN_LANTEIRA_L2",
"PISYL_PORTUGOS_H","PISYL_PORTUGOS_L","PISYL_PORTUGOS_M"),
elev_cat =
c("low", 
"high", "low", "med", 
"high", "low", "med", "low2",
"high", "low", "med"))

write_csv(elev_cat, "data/categories_elevation_site_code.csv")

```

# References
