---
title: "Soil Organic Carbon"
output: workflowr::wflow_html
bibliography: references.bib
cls: ecology-letters.csl
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r pkg}
library(tidyverse)
library(sf)
library(terra)
library(tinytable)
```

# Introduction 
Soil organic carbon (*SOC*) is the main component of soil organic matter, being of vital importance for many soil processes. For instance,  SOC content correlates positively with soil fertility, playing a key role determining the physical, chemical and biological qualities of a soil [@Victoriaetal2012BenefitsSoil].  

To estimate SOC we used the [SoilGrid database](https://soilgrids.org/) that provides digital maps of different soil variables at a spatial resolution of 250 m [@Hengletal2017SoilGrids250mGlobal; @Poggioetal2021SoilGridsProducing], built from a global database of soil profiles (WoSIS, World Soil Information Service)[@Batjesetal2017WoSISProviding; @Batjesetal2020StandardisedSoil]. 


```{r, eval=FALSE, echo=FALSE}
# Several ways to download 
# Using soilgrids.org (limitation for a 2x2 degree tile)
# TO explore the tiles see the shapefile here: https://files.isric.org/soilgrids/latest/data/soc/ 
# 
p <- st_read("data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp") 
st_bbox(p)
# use the following in the web:
# lat: -3.4 to -2.6
# long: 36.9 to 37.2
# generated url: 
# https://maps.isric.org/mapserv?map=/map/ocs.map&SERVICE=WCS&VERSION=2.0.1&REQUEST=GetCoverage&COVERAGEID=ocs_0-30cm_mean&FORMAT=image/tiff&SUBSET=long(-3.4000,-2.6000)&SUBSET=lat(36.9000,37.2000)&SUBSETTINGCRS=http://www.opengis.net/def/crs/EPSG/0/4326&OUTPUTCRS=http://www.opengis.net/def/crs/EPSG/0/4326
```

## Extract data 

For each site we extracted the Soil Organic Carbon stock from 0-30 cm deepth (ton/ha)

```{r}
# read geoinfo to get site codes
geo <- st_read("data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp", quiet=TRUE)
geo <- vect(geo)

geodf <- geo |> as.data.frame() |>
  rownames_to_column(var = "ID") |>
  mutate(ID = as.numeric(ID)) |> 
  dplyr::select(ID, site_code, sp_code, elev_code)

soc <- terra::rast("data/raw/soc_stock_ton_ha.tif")

soc_sites <- as.data.frame(terra::extract(soc, geo, bind = TRUE)) |> 
  dplyr::select(site_code, sp_code, elev_code, soc_stock_ton_ha) |> 
  arrange(
    factor(sp_code, levels = c("halepensis", "pinaster", "nigra", "sylvestris")), 
    factor(elev_code, levels = c("low2", "low", "medium", "high")))

write_csv(soc_sites, "data/dendroadaptamed_soc.csv")
```

```{r}
soc_sites |> 
   dplyr::select(-sp_code) |> 
  rename('Soil Organic Content stock (ton ha$^{-1}$)' = soc_stock_ton_ha) |> 
  tt() |> 
  group_tt(
    i = list(
      "halepensis" = 1,
      "pinaster" = 4, 
      "nigra" = 8, 
      "sylvestris" = 11
    ))
```


# References 
