---
title: "Topography from sites"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


```{r pkg}
library(tidyverse)
library(sf)
library(terra)
library(DT)
```

```{r}
# read geoinfo to get site codes
geo <- st_read("data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp", quiet=TRUE)
geo <- vect(geo)

geodf <- geo |> as.data.frame() |>
  rownames_to_column(var = "ID") |>
  mutate(ID = as.numeric(ID)) |> 
  dplyr::select(ID, site_code, sp_code, elev_code)
```


```{r, eval=FALSE}
# read mde high precission
demsn <- rast("/Volumes/Backup/CARTOTECA_V2/MDEs/mde_sn_area_influencia.asc")

# Set projection
crs(demsn) <- "+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

# Get slope and aspect 
slopesn10 <- terrain(demsn, v="slope", unit = 'degrees')
aspectsn10 <- terrain(demsn, v="aspect", unit = 'degrees')

elev <- terra::extract(demsn, geo, xy = TRUE)
slope <- terra::extract(slopesn10, geo, xy = TRUE)
aspect <- terra::extract(aspectsn10, geo, xy = TRUE)

topo <- geodf |> 
  inner_join(elev) |> 
  inner_join(slope) |> 
  inner_join(aspect) |> 
  rename(elev = mde_sn_area_influencia) |> 
  dplyr::select(-x,-y) |> 
  mutate(elev = round(elev), 
         slope = round(slope, 1), 
         aspect = round(aspect, 1), 
         aspectFactor = cut(aspect, 
                            breaks= c(0,22.5, 67.5, 112.5, 157.5, 202.5, 247.5, 292.5, 337.5, 359.5), labels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW","N")))

write_csv(topo, file = "data/topo_sites.csv")
```


## Explore topography for sites

```{r}
topo <- read_csv("data/topo_sites.csv")

topo |> DT::datatable(options =
              list(
                autoWidth = TRUE,
                pageLength = 15), 
            filter = list(position = 'top', clear = FALSE)
            
            )
```

## Compute distance among sites 

```{r}
geopinos <- st_read("data/geoinfo/geoinfo_dendro_adaptamed_pinus.shp", quiet=TRUE)

# We have to get the same ID name as the st_distance ouput 
points <- geodf$ID

# Compute the distance matrix 
point_distances <- st_distance(geopinos)

# Create an empty dataframe to store the results
distance_df <- data.frame(point1 = character(), 
                          point2 = character(), 
                          distance = numeric())

# Loop through the points and populate the dataframe
for (point1 in points) {
  for (point2 in points) {
    distance <- point_distances[point1, point2]

    distance_df <- rbind(distance_df, 
                         data.frame(point1 = point1, 
                                    point2 = point2, 
                                    distance = distance))
  }
}

write_csv(distance_df, "data/distance_among_sites.csv")
```

```{r}
g <- geodf |> dplyr::select(ID, site_code)

distance_df |> 
  mutate(distance = round(distance, 0)) |>
  left_join(g, by = c("point1" = "ID")) |>
  rename(point1_code = site_code) |>
  left_join(g, by = c("point2" = "ID")) |>
  rename(point2_code = site_code) |> 
  relocate(point1_code, point2_code) |> 
  dplyr::select(-point1, -point2) |> 
  DT::datatable(options =
              list(
                autoWidth = TRUE,
                pageLength = 13), 
            filter = list(position = 'top', clear = FALSE)
            )
```


```{r}
min_distances <- distance_df |>
  mutate(distance = as.numeric(distance)) |>
  filter(distance > 0) |>
  group_by(point1) |>
  slice_min(order_by = distance, n = 1) |>
  ungroup()

min_distances_codes <- min_distances |>
  left_join(g, by = c("point1" = "ID")) |>
  rename(point1_code = site_code) |>
  left_join(g, by = c("point2" = "ID")) |>
  rename(point2_code = site_code)

write_csv(min_distances_codes, "data/distance_among_sites_minimum.csv")
```

```{r}
min_distances_codes |> 
  mutate(distance = round(distance, 0)) |>
  relocate(point1_code, point2_code) |> 
  dplyr::select(-point1, -point2) |> 
  DT::datatable(options =
              list(
                autoWidth = TRUE,
                pageLength = 13), 
            filter = list(position = 'top', clear = FALSE)
            )
```


## Explore spatial dependency among sites

- We want to explore the spatial distribution of the sites and the potential bias (spatial dependency) among the closest sites. 
- We downloaded an image of MODIS MOD17A3HGF from [Earth data explorer](https://search.earthdata.nasa.gov/search)
- Select the h17v05 tile (corresponding to our area of interest)
- Plot to see each pixel (500 x 500 m) and then overlay the field sites

```{r}
# Read the hdf file 
m <- terra::rast("data/raw/MOD17A3HGF.A2021001.h17v05.061.2022020140519.hdf") |> 
  project("EPSG:23030")

# select only a layer 
npp <- m$Npp_500m

# function to get bbox by sp 
bbox_sp <- function(shapefile, sp) { 
  out <- shapefile |> 
    filter(sp_code %in% sp) |> 
    st_transform(crs = 23030) |>
    st_bbox() |> 
    st_as_sfc() |> 
    st_buffer(3000, joinStyle = "MITRE", mitreLimit = 500) 
  return(out)
  }

# compute bbox
nig <- bbox_sp(geopinos, c("nigra", "pinaster"))
syl <- bbox_sp(geopinos, c("sylvestris"))
hal <- bbox_sp(geopinos, c("halepensis"))

# for annotations of plots 
geopinos_t <- st_transform(geopinos, 23030) |> 
  unite("sp_elev", c(sp_code, elev_code), remove = FALSE) 
geopinos_coords <- st_coordinates(geopinos_t)
```

### *P. nigra* and *P. pinaster*

```{r}
# nigra + pinaster 
ras_nig <- terra::crop(npp, nig)
pols_nig <- ras_nig |> terra::as.polygons(aggregate = FALSE)

plot(ras_nig, main = "pinaster, nigra")
plot(pols_nig, add=TRUE)
plot(st_geometry(geopinos_t), add = TRUE, pch = 20, col = "blue")
text(geopinos_coords[, "X"], geopinos_coords[, "Y"], labels = geopinos_t$sp_elev, col = "blue", cex = 0.8, adj = c(1.1, 1))
```

### *P. sylvestris*

```{r}
# sylvestris
ras_syl <- terra::crop(npp, syl)
pols_syl <- ras_syl |> terra::as.polygons(aggregate = FALSE)

plot(ras_syl, main = "sylvestris")
plot(pols_syl, add=TRUE)
plot(st_geometry(geopinos_t), add = TRUE, pch = 20, col = "blue")
text(geopinos_coords[, "X"], geopinos_coords[, "Y"], labels = geopinos_t$sp_elev, col = "blue", cex = 0.8, adj = c(1.1, 1))
```

### *P. sylvestris*

```{r}
# halepensis
ras_hal <- terra::crop(npp, hal)
pols_hal <- ras_hal |> terra::as.polygons(aggregate = FALSE)

plot(ras_hal, main = "halepensis")
plot(pols_hal, add=TRUE)
plot(st_geometry(geopinos_t), add = TRUE, pch = 20, col = "blue")
text(geopinos_coords[, "X"], geopinos_coords[, "Y"], labels = geopinos_t$sp_elev, col = "blue", cex = 0.8, adj = c(1.1, 1))
```







