---
title: "Drought analysis"
output: workflowr::wflow_html
bibliography: references.bib
cls: ecology-letters.csl
editor_options:
  chunk_output_type: console
---

## Introduction

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10, fig.height = 7
)
```

## Introduction 
- Assessment of the drought indices using SPEI Data since. 
- Two indices were used: SPEI-6 months and SPEI-12 months. Different temporal interval could be computed. According to [@VicenteSerranoetal2010MultiscalarDrought] short time scales are mainly related to soil-water content and river discharge in headwater areas, medium time scales are related to reservoir storages and discharge in the medium course of the rivers, and longtime scales are related to variations in groundwater storage.

```{r}
library(tidyverse)
library(purrr)
library(confintr)
library(patchwork)
```


```{r}
spei_raw <- read_csv(file = "data/dendroadaptamed_spei.csv")

spei_data  <- spei_raw |> 
  dplyr::select(date, 
                spei01 = spei_1,
                spei03 = spei_3,
                spei06 = spei_6, 
                spei09 = spei_9,
                spei12 = spei_12,
                spei24 = spei_24,
                site_code) |> 
  pivot_longer(cols = spei01:spei24, names_to = "spei") |> 
  filter(!is.na(value))
  
```


## Questions 

**Do the studied forest stands suffered similar drought conditions since 1960?**

Our hypothesis are: 

- There is not differences between forest stands located in a elevational gradient in the same area. 
- The drought conditions were different among the forest stand located in contrasting slopes 


## Analysis

### Exploratory analysis 
- Plot 

### Correlation analysis

- We performed a bootstrapped Pearson Correlation 

To compare differences of drought conditions among forest stands, we performed Pearson correlation among temporal series of SPEI-06 with bootstrapped confidence intervals. 

```{r}
spei06 <- spei_data |> 
  filter(spei == "spei06") |> 
  na.omit() |> 
  dplyr::select(date, value, site_code) |> 
  pivot_wider(names_from = site_code, values_from = value) 
```


```{r}
custom_bci <- function(data, site1, site2, nrep) {
  result <- confintr::ci_cor(data[c(site1, site2)], 
                             method = "spearman", 
                             type = "bootstrap", R = nrep)

  cor_value <- result$estimate[1]
  ci <- result$interval
  vars <- dimnames(result$estimate)

  cor_df <- data.frame(
    v1 = vars[[1]],
    v2 = vars[[2]],
    correlation = cor_value,
    ci_lower = ci[1],
    ci_upper = ci[2]
  )
  return(cor_df)
}

```


```{r, eval=FALSE}

# improve use possibly purr 
set.seed(1234)

sites <- unique(spei_raw$site_code) 

# Create all possible pairs of sites
sites_pairs <- combn(sites, 2, simplify = FALSE)

sites_pairs1 <- sites_pairs[1:68]
sites_pairs2 <- sites_pairs[70:78]

column_pairs <- c(site_pairs1, sites_pairs2)

# Apply custom function using purrr::map_df
correlation_sites1 <- purrr::map_df(sites_pairs1, ~ custom_bci(spei06, .[[1]], .[[2]], nrep = 5000))

correlation_sites2 <- purrr::map_df(sites_pairs2, ~ custom_bci(spei06, .[[1]], .[[2]], nrep = 5000))


correlation_sites <- bind_rows(correlation_sites1, correlation_sites2)
write_csv(correlation_sites, file = "data/bootstraped_cor_spei.csv")
```



```{r}
correlation_sites |> 
  unite("cors", v1:v2, remove = FALSE, sep = " vs ") |> 
  ggplot(aes(x=cors, y=correlation)) + 
 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  coord_flip() +
  geom_point() + 
  theme_minimal()
```


```{r}
library(ggpmisc)
library(cowplot)
library(patchwork)
sites_pairs <- combn(sites, 2, simplify = FALSE)

sites_pairs1 <- sites_pairs[1:68]
sites_pairs2 <- sites_pairs[70:78]

column_pairs <- c(site_pairs1, sites_pairs2)




plot_list <- list()
ylab <- "SPEI-06"
# Loop through each column pair
for (i in 1:length(column_pairs)) {
  column_pair <- column_pairs[[i]]
  
  # Select the relevant columns and pivot them longer
  data <- spei06 %>%
    select(date, column_pair) %>%
    pivot_longer(-date)
  
  # Select the Boostrap correlation
  bc <- final_result |> 
  filter(v1 == column_pair[1]) |> 
  filter(v2 == column_pair[2]) |> 
  pull(correlation) |> round(3)
  
  # Create a ggplot line plot for the current column pair
  plot <- ggplot(data, aes(date, y=value, group=name, colour=name)) +
    geom_line() +
    labs(title = paste0(
      paste(column_pair[1], "-", column_pair[2]),
      " (", bc, ")"), 
      y = ylab, x ="") +
    scale_color_manual(values=c("blue", "black")) +
    theme_minimal() +
    theme(
      legend.position = "top", 
      legend.title = element_blank()
    )
  
  # Add the plot to the list
  plot_list[[i]] <- plot
}

# Combine all individual plots using patchwork
combined_plot <- wrap_plots(plotlist = plot_list, ncol = 8)

# Display the combined plot
# print(combined_plot)

ggsave(
  combined_plot, 
  file = "output/plot_boostrap.png", 
  dpi = 300, 
  width = 30, height = 30
)
```




Boostraped Pearson correlation between series (5000 repetitions) were carried out and are indicated in bracket. 

### 


```{r}
spei_raw <- read_csv(file = "data/dendroadaptamed_spei.csv")

s <- spei_raw |> 
  dplyr::select(date, 
                spei01 = spei_1,
                spei03 = spei_3,
                spei06 = spei_6, 
                spei09 = spei_9,
                spei12 = spei_12,
                spei24 = spei_24,
                site_code) |> 
  pivot_longer(cols = spei01:spei_24, names_to = "spei") |> 
  filter(!is.na(value))
  
```


```{r}


x_spei <- function(df, spei_var = "spei", 
                   spei_indices, site_var, sites) { 
  output <- df |> 
    filter(get({{spei_var}}) %in% spei_indices) |> 
    filter(get({{site_var}}) %in% sites)
  return(output)
  }

fff <- x_spei(df = s,  
       spei_indices = c("spei_06"), 
       site_var = "site_code", 
       sites = "PISYL_PORTUGOS_L")



plot_spei <- function(df, 
                      spei_var = "spei",
                      spei_indices, 
                      site_var, 
                      sites, 
                      value_var = "value",
                      title = "",
                      ...) {
  
  # Check if columns exist in the dataframe
  if (!(spei_var %in% colnames(df) && site_var %in% colnames(df) && value_var %in% colnames(df))) {
    stop("Columns specified do not exist in the dataframe.")
  }
  
  spei_selected <- df %>%
    filter({{spei_var}} %in% spei_indices,
           {{site_var}} %in% sites,
           !is.na({{value_var}})) %>%
    mutate(
      sign = ifelse({{value_var}} > 0, "pos", "neg"),
      variable = recode({{spei_var}},
                        "spei01" = "1-month",
                        "spei03" = "3-month",
                        "spei06" = "6-month",
                        "spei09" = "9-month",
                        "spei12" = "12-month",
                        "spei24" = "24-month",
                        "spei48" = "48-month"
      )
    )
  
  # p <- ggplot(data = spei_selected,
  #             aes(x = .data[[date]],
  #                 y = .data[[value_var]], 
  #                 fill = .data[[sign]], 
  #                 colour = .data[[sign]])) +
  #   geom_bar(stat = "identity") +
  #   facet_grid(.data[[site_var]] ~ .data[[variable]]) + 
  #   scale_fill_manual(values = c("pos" = "lightblue", "neg" = "tomato")) +
  #   scale_colour_manual(values = c("pos" = "lightblue", "neg" = "tomato")) +
  #   theme_bw() +
  #   labs(
  #     title = title,
  #     x = "",
  #     y = "SPEI"
  #   ) +
  #   theme(
  #     strip.background = element_rect(fill = "white"),
  #     legend.position = "none",
  #     ...
  #   )
  # 
  return(spei_selected)
}



kk <- plot_spei(df = s, spei_indices = "spei_06", 
                site_var = "site_code", sites = "PISYL_PORTUGOS_L")


plot_spei <- function(df, 
                      spei_var = "spei",
                      spei_indices, 
                      site_var, 
                      sites,
                      title, ...) {
  
  spei_selected <- df |>
    filter(get({{spei_var}}) %in% spei_indices) |> 
    filter(get({{site_var}}) %in% sites) |> 
    # filter(!is.na({{value}})) |>
    mutate(
      sign = ifelse(.data[[value]] > 0, "pos", "neg"),
      variable = recode({{spei_var}},
                        "spei01" = "1-month",
                        "spei03" = "3-month",
                        "spei06" = "6-month",
                        "spei09" = "9-month",
                        "spei12" = "12-month",
                        "spei24" = "24-month",
                        "spei48" = "48-month"
      )
    )
  # 
  p <- ggplot(spei_selected,
              aes(x = .data[[date]],
                  y = .data[[value]], fill = .data[[sign]], colour = .data[[sign]])) +
    geom_bar(stat = "identity") +
    facet_grid(.data[[site_var]]~.data[[variable]]) +
    scale_fill_manual(values = c("pos" = "lightblue", "neg" = "tomato")) +
    scale_colour_manual(values = c("pos" = "lightblue", "neg" = "tomato")) +
    theme_bw() +
    labs(
      title = title,
      x = "",
      y = "SPEI"
    ) +
    theme(
      strip.background = element_rect(fill = "white"),
      legend.position = "none"
    )

  return(spei_selected)
}


plot_spei(df = s, spei_indices = "spei_06", site_var = "site_code", sites = "PISYL_PORTUGOS_L")





```



# References 
